\documentclass{article}
\usepackage{fullpage}
\usepackage{times}
\usepackage{bibunits}
\usepackage{url}
\usepackage[ManyBibs,NoDate]{currvita}




\textwidth 6.6in
\oddsidemargin -.05in
\evensidemargin -.05in
\textheight 9.2in
\topmargin -.1in

\newcounter{bibctr}

\newcommand{\mkbib}[1]{
    \refstepcounter{bibctr}
    \item[\hskip0pt plus1filll \hbox{[\arabic{bibctr}]}]
  \begin{bibunit}[plain]
      \nocite{#1}
      \putbib[ref,rfc,talks]
    \label{bib:#1}
  \end{bibunit}
}

\newcommand{\mkbiba}[2]{
    \refstepcounter{bibctr}
    \item[\hskip0pt plus1filll \hbox{[\arabic{bibctr}]}]
  \begin{bibunit}[plain]
      \nocite{#1}
      \putbib[ref,rfc,talks]
    \label{bib:#1} \vspace{-0.05in}
    {\em Acceptance rate: #2\%}
  \end{bibunit}
}


\newcommand{\refbib}[1]{[\ref{bib:#1}]}

\newcommand{\ie}{{\em i.e.}}
\newcommand{\eg}{{\em e.g.}}

\def\up#1{\raise.16ex\hbox{#1}}

\begin{document}
\begin{cv}{}

\noindent {\nf Nick Feamster}\\

\begin{tabular}{p{3.25in}p{4in}}
\noindent 
Professor & \url{feamster@cs.princeton.edu} \\
Department of Computer Science & \url{http://www.cs.princeton.edu/~feamster/} \\
Princeton University  & \url{http://connectionmanagement.org/}\\
35 Olden Street &  \url{http://noise-lab.net/}\\
Princeton, NJ 08540 \\
\end{tabular}


\subsection*{Paper and Publication Awards}
\begin{itemize}
\itemsep=-1pt
\item ACM SIGCOMM Community Award, ACM SIGCOMM Internet Measurement
  Conference (2013)
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal

\subsection*{Journal Publications}
\label{subsec:journals}

\begin{pub}
\mkbib{feamster2014:ccr}
\mkbib{burnett2013:censorship}
\mkbib{kim2013:improving}
\end{pub}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Conference

\subsection*{Conference Publications}


% \hspace*{-0.75in}{\it\large Internet Routing} 

\begin{pub}
\mkbiba{kim2015:nsdi}{19}
\mkbiba{chetty2015:chi}{23}
\mkbiba{sundaresan2015:pam}{27}
\mkbib{chavula2014:africomm}
\mkbiba{narain2014:ccs}{19}
\mkbiba{jones2014:imc}{22}
\mkbiba{gupta2014:sigcomm}{19}
\mkbiba{sundaresan2014:atc}{15}

\mkbiba{gupta2014:pam}{31}
\mkbiba{xing2014:pam}{31}

\mkbiba{chetty2013:dev}{33}
\mkbiba{mundada2013:silverline}{19}

\mkbiba{Hao2013:dns}{23}
\mkbiba{Grover2013:nat}{23}
\mkbiba{Sundaresan2013:webperf}{23}
\vspace*{-0.05in}
{\bf ACM SIGCOMM Community Contribution Award.}


\mkbiba{Xing2013:poison}{16}
\mkbiba{Valancius2013:pecan}{11}
\end{pub}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Workshops


\subsection*{Workshop Publications}

\begin{pub}

\mkbib{schinkler2014:hotnets}
\mkbib{donovan2014:hotnets}
\mkbib{jones2014:foci}

\mkbib{feamster2013:isoc-latency}
\mkbib{chetty2013:isoc-latency}

\mkbiba{anwer2013:slick}{23}
\mkbiba{feamster2013:sdx}{23}
\end{pub}


\section*{Research Proposals and Grants}
\label{subsec:research}

\def\funding#1#2#3#4#5{\item{{{\bf #1}}
\newline Sponsor: {#2}
\newline Investigator(s): {#3}
\newline Amount: {\sl #4}
\newline {\sl #5}}}

\begin{grants}

\funding{A Software Defined Internet Exchange}{National
  Science Foundation}{N. Feamster (PI),
  J. Rexford, S. Shenker}
{\$500,000 for 3 years.}{Awarded: August 2014} 

\funding{Studying and Improving the Performance of Access
  Networks}{National
  Science Foundation}{N. Feamster (PI), A. Snoeren}
{\$281,364 for 3 years.}{Awarded: August 2014} 

\funding{EPICA: Empowering People to Overcomme Information Controls and
  Attacks}{National
  Science Foundation}{W. Lee (PI), N. Feamster, H. Klein, M. Bailey, M. Chetty}
{\$750,000 for 4 years.}{Awarded: August 2014} 

\funding{An Open Platform for Internet Routing Experiments}{National
  Science Foundation}{N. Feamster, E. Katz-Bassett, D. Choffnes 
  (PI)}{\$361,617 for 3 years.}{Awarded: August 2014} 

\funding{An Open Observatory for the Internet's Last Mile}{National
  Science Foundation}{N. Feamster
  (PI), S. Banerjee, J. Cappos}{\$399,879 for 3 years.}{Awarded: June 2014} 

\funding{Improving the Performance and Security of Home Networks with
  Programmable Home Routers}{Comcast}{N. Feamster (PI)}{\$65,000 for 1
  year}{Awarded: September 2013} 

\funding{Demand Characterization and Management for Access Networks}{Cisco
  Systems}{N. Feamster (PI) and R. Johari}{\$99,766 for 1 year}{Awarded:
  April 2013}

\end{grants}


\section*{Seminar Presentations}
\label{subsec:seminar}

%\subsection{Talks}

\begin{pub}

% December 2014
% GAIA
\mkbib{gaia:peering:2014}

% November 2014
% AT&T, DC-APS
\mkbib{att:sdx:2014}
\mkbib{dcaps:encore:2014}

% October 2014
% BITAG, WebTAP
\mkbib{bitag:sdx:2014}
\mkbib{webtap:bobble:2014}

% September 2014
% NSA
\mkbib{nsa:sdn:2014}

% August 2014
% MOOC 
\mkbib{mooc:2014}

% July 2014
% Citizen Lab Summer Institute
\mkbib{clsi:censorship:2014}

% May 2014
% ONUG
\mkbib{onug:sdn:2014}

% April 2014
% Skytree, SDN Meetup (DC), Princeton
\mkbib{skytree:ml:2014}
\mkbib{princeton:bots:2014}
\mkbib{sdn:meetup:2014}

% March 2014
% AIMS, DIMACS
\mkbib{aims:bismark:2014}
\mkbib{dimacs:silverline:2014}

% December 2013
\mkbib{columbia:censor:2013}
\mkbib{nitrd:sdx:2013}
\mkbib{conext:student:2013}

% November 2013
\mkbib{princeton:censor:2013}

% October 2013
\mkbib{sdn:gfi:2013}
\mkbib{censorship:zisc:2013}

% September 2013
\mkbib{juniper:sdn:2013}

% August 2013
\mkbib{cisco:sdi:2013}

% July 2013
\mkbib{mooc2013}

% June 2013
\mkbib{bismark:afnog:2013}

% May 2013
\mkbib{infocontrol:georgetown:2013}

% April 2013
\mkbib{sdx:ons2013}
\mkbib{bismark:nae:2013}

% March 2013
\mkbib{sdx:irtf2013}
\mkbib{infocontrol:berkeley:2013}
\mkbib{bobble:aims:2013}
\end{pub}


\section*{Department and University Service}

(All service listed below was at Georgia Tech, where I was on the
faculty through December 2014.)

\subsection*{Department and University Committees}

\begin{itemize}
\itemsep=-1pt
\item Faculty Recruiting Committee (2013)
\item Online Masters in Computer Science -- Founding Committee (2013--2014)
\item Ph.D. Curriculum Committee (2013--2014)
\item Organizer for Ph.D. Student Visit Day (2013)
\end{itemize}

\subsection*{Informal Outreach}

\begin{itemize}
\itemsep=-1pt
\item Serving on the Broadband Internet Technical Advisory Group (BITAG).
  Published a technical report on interconnection and peering that
  was presented to the Federal Communications Commission (FCC) and
  several congressional committees. (Fall 2014)
\item Founded and maintain a research advice blog,
  \url{http://greatresearch.org/}, that receives thousands of reads
  every month. (Fall 2013--Present)
\item Serving as a host institution for the Open Technology Fund's (OTF)
  Information Controls Fellowship Program. (Fall 2014--Present)
\item Obtained ethics/responsible conduct of research (RCR)
  certification for the ``Intro to the Ph.D.'' course, which I designed
  and have taught for about eight years. (Fall 2013)
\item Informally advising a student at University of Cape Town, Josiah Chavula, on
  networking research. (Summer 2014--Present)
\item Hosted a ``meetup group'' on Software Defined Networking at
  Georgia Tech for presentations and discussions of SDN technologies in
  the Atlanta area. (September 2014)
\item Panelist in School of Computer Science on academic job search advice
  (Fall 2014).
\item Meeting with prospective undergraduates to discuss the
  undergraduate CS program at Georgia Tech. (March 2014)
\item Gave talks on Massive Open Online Courses (MOOCs) and Georgia
  Tech's Online MS CS program (OMS~CS) at various universities,
  including MIT and the University of Cape Town. (Summer 2014)
\item Speaker at Undergraduate Research Opportunity Program symposium
  (Fall 2013, Fall 2014).
\end{itemize}

\end{cv}

\end{document}
