\documentclass[letterpaper,11pt]{article}

  \usepackage[hmargin=.65in,vmargin=.6in,headsep=.2in,centering]{geometry}

  \usepackage{multirow}

  \pagestyle{plain}

  \newcommand{\onerow}[1]{%
    \multirow{3}{*}{#1}%
      & %
      & %
      & %
      & %
      & %
      \\[16mm]\hline%
  }

  \renewcommand{\arraystretch}{2}

\begin{document}

  \hspace{\fill}Name: \underline{\hspace{3in}}

  \vspace{.15in}

  \begin{center}

    \Large CMSC 330 Office Hours Preferences

  \end{center}

  \vspace{3mm}

  \enlargethispage{4mm}

  \begin{tabular}{|r|*{5}{p{1.15in}|}|}

     \hline

     % empty
       & \multicolumn{1}{|c|}{Monday}
       & \multicolumn{1}{c|}{Tuesday}
       & \multicolumn{1}{c|}{Wednesday}
       & \multicolumn{1}{c|}{Thursday}
       & \multicolumn{1}{c|}{Friday}
       \\ \hline \hline

    \onerow{10}

    \onerow{11}

    \onerow{12}

    \onerow{1}

    \onerow{2}

    \onerow{3}

    \onerow{4}

    \onerow{5}

  \end{tabular}

  \bigskip

  \begin{center}

    (\emph{\LARGE instructions on reverse})

  \end{center}

  \pagebreak

  \begin{itemize}

    \item On the chart on the other side, please put an ``X'' through all
          times you \textsc{absolutely cannot} have office hours (you have
          class of your own during those times, you have a religious
          obligation, a research meeting, etc.).  If you're not certain
          about your class schedule yet, put an X through the times of all
          classes you \emph{might} be taking.

    \item Teaching TAs, put an ``X'' through the times you will be
          conducting discussion sections.

    \item Mark all of the hours that would be your preferred choice for
          office hours with a 1.  Mark as many of the hours that would work
          out well for your schedule to have office hours with a 1 as
          possible.  We'll pick the right number of these hours for each TA
          and inform you later of the schedule via email.

          The more hours you mark as \#1 preferred hours, the greater the
          chance we can assign office hours for you that you would want.
          If you were to mark only a few hours as \#1 preferred hours, it's
          very doubtful we would be able to give you only those exact
          hours.

    \item Mark any second choice office hours that would be acceptable but
          not your preferred times with a ``2''.

          Teaching TAs, don't mark the times when your students will be in
          lecture as preferred office hours times (with a 1).  If you do,
          your own students won't be able to come to your office hours.  We
          can let you know when your students will be in lecture, or just
          check the online schedule of classes on Testudo.  You can mark the
          times when your students will be in lecture as \#2 if you wouldn't
          mind having office hours then, but don't mark them as \#1.  We'll
          try to avoid giving \#2 hours, but usually TAs end up some \#1
          hours and some \#2 hours.

    \item You can draw lines and mark desired hours by the half-hour, or
          even by 15 minutes if needed.

    \item In marking possible office hours, take your class times and
          locations into account.  If you have class beginning or ending at
          a certain time (especially if it's in a building far away from
          this one) then don't indicate preferred office hours immediately
          before or after your class; leave yourself some time to return to
          the AVW building or to get to class before or after your preferred
          office hours time.  For instance, if you have class starting at
          2:00.  don't say that you can have office hours up until 2:00,
          because you would have to leave your office hours early to get to
          class on time.  Instead indicate that you can have office hours
          until 1:30 or 1:45.

    \item Be sure to fill out \textbf{\underline{all the entries}} in the
          table.  Every time should have either some number or an X.

    \item If you have other special preferences (e.g., something like ``I
          prefer to have office hours later in the day'') please make a note
          of them; we'll try to accomodate them if possible.

  \end{itemize}

  \noindent
  \textbf{If anything changes in your schedule in the next few days, inform
  us immediately.}

  \medskip

  \noindent
  \textbf{Once the office hours schedule is made final and given out to
  students, TAs may not change their schedules such that the office hours
  schedule would have to be altered.}

  \bigskip\medskip

  \renewcommand{\arraystretch}{1.5}

  \begin{center}

    {\Large Please provide the following information:}

  \end{center}

  \vspace{-2mm}

  \noindent
  \begin{minipage}[t]{\textwidth}

    \renewcommand{\arraystretch}{2.25}

    \begin{tabular}{p{4.1in}@{\hspace{4mm}}l}

      Phone number:
        & \underline{\hspace{2.5in}}
        \\

      Office room number, if you have an office in AVW:
        & \underline{\hspace{2.5in}}
        \\

      If you have another phone number where you may be reached please let
      us know it:
        & \underline{\hspace{2.5in}}
        \\

    \end{tabular}

    \medskip

    (If you don't know all of this information yet, please let us know when
  you get it.)

  \end{minipage}

\end{document}
