\documentclass{article}
\usepackage{fullpage}
\usepackage{times}
\usepackage{url}
\usepackage{fancyhdr}

\newcommand{\ie}{{\em i.e.}}
\newcommand{\eg}{{\em e.g.}}

\newcommand{\fp}{\vspace*{0.05in}\noindent}

\renewcommand{\paragraph}[1]{\fp {\bf #1}}


\newpage \setcounter{page}{1}
\lfoot[Printed: \today]{Nick Feamster}
\cfoot{\fancyplain{}{Revised: December 21, 2013}}
\rfoot{Page \thepage~of 2}

\thispagestyle{empty}

\begin{document}

\vspace*{-0.45in}
\begin{center}
{\Large\textbf{Nick Feamster}}\\[0.1in] {\large\textbf{
Teaching Statement}}\\
%feamster@cc.gatech.edu \\
%http://www.cc.gatech.edu/$\sim$feamster/ \\[.1in] 
\end{center}
I believe that students learn best by doing.  Hands-on experience and
real-world exercises not only make classes more exciting, but they
also provide memorable examples and analogies.  Our ability to
understand abstract concepts is more limited than our ability to process
concrete examples that relate to familiar ideas.  Some of
my own memorable classroom experiences were lectures with concrete
examples; I apply a similar approach in my courses, tying abstract
concepts to relevant concrete examples and hands-on experience.  For
example, ``route hijacking'' is more concrete when students can
actually see network traffic going where it isn't supposed to go.
%Routing protocols make more sense when students can configure routers,
%induce failures on an experimental network, and watch the behavior of
%traffic over that network.

To this end, my first goal in teaching both networking and security
classes is to connect textbook material with (1)~real-world examples;
and (2)~hands-on experience.  At both the undergraduate and graduate
levels, I have incorporated course material that familiarizes students
with the state of the art in network design, implementation, and
experimentation; for example, I have students run experiments on network
testbeds like Emulab, VINI, and BISmark, with real software routers that
they can configure.  Where appropriate, I have also allowed students to
shape the course themselves by bringing in real-world examples, from
which I will design a lecture.  For example, in my undergraduate
networking course, I maintain a wiki where students could post relevant
``current events'' in networking and vote on which topics they would
like to see covered.  Based on that input, I incorporate new material
into the syllabus---teaching not just the current event itself, but also
the relevant foundational material.  I also bring my own current events
to lecture and relate them to the textbook being covered.  This takes
more effort than dusting off old notes, but it keeps students engaged
and helps me stay abreast of what is happening both in industry and in
research.

My second goal is to elevate course material so that students are not
just learning mechanics of protocols and systems, but also gaining a
deeper understanding of the {\em concepts} that underlie their design.
With so many computing and communications tools for aggregating and
processing information, conventional lectures are no longer always the
most efficient way to convey textbook information.  In my lectures, I
try to go beyond what is taught in the textbook---rather than teaching
only mechanics, I ask students about design rationales, and whether they
would make the same choices today, given the changing roles of
communications networks.  Given that networking is still maturing as a
field, there is a tendency to focus on protocol details, which may
change over time.  

\paragraph{Teaching research methods.} I have designed a graduate
course that focuses on teaching first-year graduate students how to do
research.  The course includes topics ranging from paper reading to
fellowship applications to generating (and executing) research
ideas. The value of the course is evident from student response:
Students at various points along their Ph.D.---even more senior
students---sit in on lectures.  We presented a paper on the course at
{\em SIGCSE} in 2008, and other computer science departments are
adopting the material. I also have developed a blog on research
methods to accompany the course (\url{http://greatresearch.org/}),
which receives about 10,000 page views every month. 

\paragraph{Bringing new media and technology into the classroom.}
I believe that traditional lectures are ``going the way of the
blackboard'', and I have made concerted efforts to introduce new media
and methods into the classroom.  I have now designed two massive open
online courses (MOOCs), and I am using technology such as video lectures
in the classroom to create a ``flipped classroom'' experience.  I have
also begun to disseminate the results of my research and teaching
efforts through online media that can reach a much greater set of
people.  I have developed two blogs---one about my own research and one
about research methods.  One of my blog entries on Internet censorship
was picked up by {\em The Wall Street Journal}; the article won the
Bronze Anvil journalism award.  I have also developed the first Massive
Open Online Course (MOOC) on Software Defined Networking, which had
50,000 enrolled students in the first offering, 1,000 of whom
successfully completed all of the programming assignments.  I used the
videos from the MOOC as part of a ``flipped classroom'' seminar for
students at Georgia Tech in Fall 2013.

\paragraph{Mentoring and advising.}
My graduate students have received best paper awards at top conferences
(including {\em SIGCOMM}), and my undergraduate students I have advised
have garnered publications in top conferences and have gone to the best
computer science graduate schools in the country.  One of my most
rewarding advising experiences was with undergraduate student Megan
Elmore.  Under my advisement, she received the Sigma Xi Undergraduate
Research Award, the CoC undergraduate research award, and an NSF
Graduate Fellowship; she is now a Ph.D. student at Stanford
University. I have received multiple ``Thank a Teacher'' awards at
Georgia Tech for outstanding teaching performance.  I also won a
Hesburgh Teaching Fellow award at Georgia Tech in 2012, which is awarded
to only about ten instructors across campus.


\end{document}
