\lfoot{\fancyplain{Printed: \today}{Irfan Aziz Essa}}
\cfoot{\fancyplain{For GT Internal USE}{Complete C.V. (Internal)}}
%\cfoot{\fancyplain{Modified: \today}{Complete C.V.}} %
\rfoot{Page  \thepage}

\maketitle

\vspace{-.3in}
\begin{small}
\tableofcontents
\end{small}
%\vfill
\newpage

\setlength{\parskip}{1pt}
\section*{EDUCATIONAL BACKGROUND}
\addcontentsline{toc}{section}{EDUCATIONAL BACKGROUND}
\label{sec:educational}
\begin{center}
\begin{tabular}{lllr}

{\bf Degree} & {\bf Year} & {\bf University}\hspace{2.0in} & {\bf Field} \\
\hline \\[\tabitemskip]

\textbf{Ph.D.} & 1995 & Massachusetts Institute of Technology, &
{\sl Media Arts \& Sciences}\\
& & Cambridge, MA., USA. \\[\tabitemskip]
\textbf{S.M.} &  1990 & Massachusetts Institute of Technology, &
{\sl Engineering}\\
& & Cambridge, MA., USA.\\[\tabitemskip]
\textbf{B.S.} & 1988 & Illinois Institute of Technology, &
{\sl Engineering} \\
& & Chicago, IL., USA.
\end{tabular}
\end{center}

\bigskip

\section*{EMPLOYMENT HISTORY}
\addcontentsline{toc}{section}{EMPLOYMENT HISTORY}
\label{sec:employment}

\begin{center}
\begin{tabular}{llr}
\textbf{Title} & \textbf{Organization} & \textbf{Years} \\ \hline \\[.1in]

\textbf{Associate} & College of Computing & {\sl 2002-present}
\\
\textbf{Professor} & Georgia Institute of Technology, \\
& Atlanta, GA., USA\\[\tabitemskip]

\textbf{Adjunct Assistant} & School of Electrical and Computer
Engineering, & {\sl 1998-present} \\
\textbf{Professor} & Georgia Institute of Technology, \\
& Atlanta, GA., USA \\[\tabitemskip]

\textbf{Assistant} & College of Computing & {\sl 1996-2002}
\\
\textbf{Professor} & Georgia Institute of Technology, \\
& Atlanta, GA., USA\\[\tabitemskip]


\textbf{Research} & Media Laboratory, Perceptual Computing
Section, & {\sl 1994-1996} \\
\textbf{Scholar/Scientist} & Massachusetts Institute of
Technology, \\
& Cambridge, MA., USA \\[\tabitemskip]

\textbf{Research} & Media Laboratory, Perceptual Computing
Section, & {\sl 1988-1994} \\
\textbf{Assistant} & Massachusetts Institute of Technology,\\
& Cambridge, MA., USA. \\[\tabitemskip]

\textbf{Research} & Vision, Graphics and Visualization Group,&
{\sl Summer 1990} \\
\textbf{Intern} & Thinking Machines Corporation, \\ & Cambridge,
MA., USA

\end{tabular}
\end{center}

%\bigskip

\section*{CURRENT FIELDS OF INTEREST}
\addcontentsline{toc}{section}{CURRENT FIELDS OF INTEREST}
\label{sec:current}

Computer Vision, Computer Graphics, Human-computer Interaction,
Machine Learning, Numerical Modeling, Computational Perception and
Artificial Intelligence.

%\bigskip

\subsubsection*{Goals:}

My research focuses on the analysis, interpretation, and synthesis
of video. In particular, I am interested in developing methods for
dynamic (and sometimes causal) analysis of the spatio-temporally
variable video signals. Such methods allow for interpretation of
what is happening in a scene and for plausible synthesis of
similar scenes. I am motivated by the desire to build machines
that can recognize complex behaviors and mimic (synthesize)
similarly complex actions by observing (seeing) them. My work
to-date has been aimed at making state-of-the-art contributions to
computer vision and computer graphics, relying heavily on
numerical modeling and learning, image processing, and control
theory. My research will have potential impact on work in
artificial intelligence and human-computer interaction.

%\begin{itemize}
%
%
%\item{To build machines that determine what is happening in a
%scene. Such machines will be aware of what people are doing and
%support the human activity appropriately.}
%
%\item{Use video data to extract realistic models of complex scenes
%and/or people. Use these models to generate perceptually correct
%synthesis of such scenes.}
%
%\end{itemize}


\newpage
\section{TEACHING}
\label{sec:teaching}

\subsection{Courses Taught}
\label{subsec:courses}

\begin{center}
\begin{tabular}{llrl}

& & \textbf{Number of \newline } &\\
\textbf{Quarter/Year} & \textbf{Course Number \& Title} &
\textbf{Students} & \textbf{Comments} \\ \hline \\[\tabitemskip]

\multicolumn{4}{c}{\textbf{Media Laboratory, Massachusetts
Institute of Technology.}}\\[\tabitemskip]

Spring, 1995 & MAS 961 Machine Understanding of Video & 15 & New\\[\tabitemskip]

\multicolumn{4}{c}{\textbf{College of Computing, Georgia
Institute of Technology.}}\\[\tabitemskip]

Winter, 1997  & CS 7321 Low-level Computer Vision & 26  & Revised\\[\tabitemskip]
Spring, 1997 & CS 7322 High-level Computer Vision & 15 & Revised\\[\tabitemskip]
Fall 1997 & CS 7100 Introduction to Grad Studies  & 40  & Revised\\[\tabitemskip]
Fall 1997 & CS 8113 Computational Perception  & 15  & New \\[\tabitemskip]
Winter 1998  & CS 7321 Low-level Computer Vision  & 15 & \\[\tabitemskip]
Spring 1998  & CS 7322 High-level Computer Vision  & 20 & \\[\tabitemskip]
Fall 1998  & CS 7100 Introduction to Grad Studies  & 41 & \\[\tabitemskip]
Spring 1999  & CS 4803/8113 Digital Special Effects  & 22  & New\\[\tabitemskip]
Spring 2000 &  CS 4480 Digital Video Special Effects &  32  &
New\\[\tabitemskip]
Spring 2001 &  CS 4480 Digital Video Special Effects &  26
&\\[\tabitemskip]
Spring 2002 &  CS 4480 Digital Video Special Effects & 25
&  \\[\tabitemskip]
Spring 2002 &  CS 4496/7496 Computer Animation & 33
& Revised \\[\tabitemskip]
\multicolumn{4}{l}{\textbf{Seminars}}\\[\tabitemskip]
Winter 97-Fall 98 & Software Agents & 7-10 & Seminar\\[\tabitemskip]
Winter-Spring, 98 & Technology and Society & 10& Seminar\\[\tabitemskip]
Fall 97-Spring 00 & Future Computing Environments &  15-25 &
Seminar


\end{tabular}
\end{center}

%\bigskip

\subsection{Curriculum Development}
\label{subsec:curriculum}

\centerline{\textbf{Media Laboratory, Massachusetts Institute of
Technology.}}

\medskip

\paragraph{MAS 961 Machine Understanding of Video:} A new graduate course
dealing with applications of computer vision and audition
techniques for automatic annotation of video. A project-based
course in which groups of 4 students each designed and implemented
systems for processing and analysis of video and audio signals.

\medskip

\centerline{\textbf{College of Computing, Georgia Institute of
Technology, Atlanta, GA.}}

%\bigskip

\paragraph{CS 7321 Low-level Computer Vision (1997-98):} A graduate
introductory course for computer vision. Revised and updated old
content. Hands-on work on different techniques of computer vision
and image understanding in the form of projects and computer labs
was added. In addition to theoretical concepts, emphasis was also
given to real world problem solving.

\paragraph{CS 7322 High-level Computer Vision (1997-98):} Continuation of
low-level computer vision concepts from CS7321. Revised and updated old
content. Emphasis laid on recognition and modeling methods for computer
vision. Hands-on work on different techniques of computer vision and image
understanding in the form of projects and computer labs was added. In
addition to theoretical concepts, emphasis was also given to real world
problem solving.

\paragraph{CS 8113 Computational Perception (1997):} Designed and taught a
completely new class aimed at exploring various issues of machine perception
of signals (video and audio) for interaction and analysis. Material covered
included study of both the engineering and psychological issues of
perception. Students worked hands-on towards the design and development of
projects, in teams, which dealt with machine perception.

\paragraph{A curriculum for Digital Imaging in Science and Engineering
Education (In Progress):} Working with faculty from the Department
of Electrical and Computer Engineering's Center for Signal and
Image Processing and the College of Computing's Graphics,
Visualization and Usability Center to develop and institute an
Undergraduate/Graduate Certificate Program for Digital Imaging in
Science and Engineering Education.

\paragraph{Seminar on Technology and Society (1998):} (with J. Berman and Dr.
T. Harpold) Started a new seminar class aimed at addressing the
issues relating to the effects of modern (computing) technology to
past, present, and future societies. A weekly seminar with
participants from different units at Georgia Tech, in the format
of round table discussions and invited speakers to bring to bear
ideas of where technology has taken us, and where it will take us.

\paragraph{Conversion to Semesters (1998-99):} In preparation for the semester
conversion, redesigned the computer vision and computational perception
courses for semesters. The two quarter series of computer vision courses was
merged into a semester long computer vision course. The computer vision
course will be offered as a 4000 level undergraduate course as well as a
7000 level graduate course. The computational perception course will be
offered as a 7000 level graduate course.

\paragraph{CS 4480 Digital Video Special Effects (1999-2002):} Developed a new
class aimed at teaching undergraduates the basics of computing
with video and audio. Students learn about digital representations
of video and study manipulation, processing, interpretation and
synthesis of video under the guise of generating special effects.
Student in groups produce videos to showcase their technical
skills of dealing with digital video, which are shown in an open
public forum. Updated the class for the semester version.
Technology Licensing funds (about \$ 150,000) were acquired to
upgrade the classrooms and to purchase audio/video equipment to
support this class.  Hewlett Packard and Intel donated computer
for teaching of this class. Received \textsl{Georgia Tech. Outstanding use of Innovative Technologies in Teaching Award, 2000} for designing and teaching of this course.

\paragraph{CS 4496/7496 Computer Animation (2002):} Redesigned and updated the Computer Animation course, and offered it for the first time after 1999.  Adapted all of the newer material that has appeared in this field in the last few years.  Emphasized traditional animation principles while covering fundamental approaches that are key for computer methods. Equal coverage of keyframing, motion capture, and simulation methods of animation.  Students (both grad and undergrad) did assigned problem sets and did a final project.

\medskip

%\bigskip

\def\student#1#2#3#4#5{\item[#1]{#2
\newline{\sl #3}
\newline Publications: {\sl #4}
\newline {\sl #5}}}

\def\studentnopub#1#2#3#4{\item[#1]{#2
\newline{\sl #3}
%\newline Publications: {\sl #4}
\newline {\sl #4}}}

\def\studentnonote#1#2#3#4{\item[#1]{#2
\newline{\sl #3}
\newline Publications: {\sl #4}}}
%\newline {\sl #4}

\def\studentempty#1#2#3{\item[#1]{#2
\newline{\sl #3}}}

\subsection{Individual Student Guidance}
\label{subsec:individual}

\subsubsection*{Post-Doctoral Fellows}

\begin{description}

\student{Lionel Reveret}{~}{Spring 2000 - Fall 2001.}{
\ref{pub:affect-nips00}, and \ref{pub:reveret01a}.} 
%and \ref{pub:reveret01b}
{Research on building Conversational
Agents. \newline \textbf{At INRIA France, IMAGIS Group in
Grenoble as \emph{Charge de recherches}}. }

\end{description}

\subsubsection*{Ph.D. Students Supervised (in process as well as
graduated)}


\begin{description}

\student{Darnell J. Moore}{(ECE, with Professor M. Hayes)}{Winter
1997 - Spring 2000.}{\ref{pub:actions-avbpa99},
\ref{pub:actions-iccv99}, \ref{pub:moore01a} and \ref{pub:moore02a}.} {Vision-Based
Recognition of Actions using Context.
\newline Member of Research Staff Texas Instrument
Corp. Dallas, TX. USA. 2000 - Present.}

\student{Arno Sch\"{o}dl}{(CoC)}{Fall 1997 - Summer 2002.} {
\ref{pub:headtrack-pui98}, \ref{pub:headtrack-pdpta99},
\ref{pub:vidtex-siggraph00}, \ref{pub:vidtex-nips00},
\ref{pub:schoedl01a}, \ref{pub:schoedl02a}, and \ref{pub:video-matting01}.}{Multi-Dimensional Exemplar-Based Texture Synthesis.
\newline \textbf{Awarded Microsoft Research Fellowship 2000-2002.}
\newline \textbf{Awarded Georgia Tech College of Computing's Outstanding Graduate Student Research Award 2002.}
\newline Founder and Technical Director, Think-Cell Software, Berlin, Germany, 2002 - Present}

\student{Gabriel J. Brostow}{(CoC)}{Fall 1997 - Present.}
{\ref{pub:dance-ijhci00}, \ref{pub:layers-iccv99},
%\ref{pub:dvfx01}, 
and \ref{pub:blur-siggraph01}.}{Current Research
on 3D tracking from video.
\newline \textbf{Awarded NSF HCI Trainee Fellowship 1998-2000.}
\newline \textbf{Awarded Intel PhD Student Fellowship 2000-2001.}
\newline \textbf{Awarded Georgia Tech College of Computing's Outstanding Graduate Student Teaching Award 2002.}}

\student{Antonio Haro}{(CoC)}{Fall 1997 - Present.}{
\ref{pub:headtrack-pui98}, \ref{pub:pupil-chi00},
\ref{pub:pupil-cvpr00}, \ref{pub:skin-eg-render01}, and \ref{pub:haro02a}.} {Current
research on facial modeling and non-photorealistic rendering.
\newline \textbf{Awarded AT\& T Graduate Fellowship 1998-2002.}}

\student{Drew Steedly}{(CoC)}{Spring 1998 - Present.} {
\ref{pub:sfm-iccv01}.} {Current Research on Optimal Structure from
Motion from Video. \newline Transferred from Georgia Tech's School
of ECE in Fall 1998. \newline \textbf{Awarded Intel PhD Student
Fellowship 2001-2002.} }

\studentnopub{Alan Chen}{(CoC)}{Fall 2000 - Present.} {Current
Research on Generating Non-photorealistic Renderings of Motion.}

\studentnopub{Daniel Sternberg}{(CoC)}{Fall 2001 - Present.} {Current
Research on Representations of Human Motion.}

\studentnopub{David Minen}{(CoC, with Professor Thad Starner)}{Fall 2001 - Present.} {Current
Research on Human Activity Recognition.
\newline \textbf{Awarded NSF Graduate Student Fellowship 2002 - 2005.}}





\end{description}

\subsubsection*{Ph.D. Special Problems students.}

%\info{Name, Dates, Topic.}

\begin{description}


\student{Andrew Gardner}{(ECE)}{Summer 1997 - Fall
1997.}{\ref{pub:affect-pui97}.}{Title: ``Audio in Interactive
Environments.''}

\studentempty{Alexander Stoychev}{(CoC)}{Fall 1998 - Summer 1999.}

\student{Scott Stillman}{(ECE)}{Winter 1997 - Spring 2002.}
{\ref{pub:tracking-avbpa99} and \ref{pub:stillman01a}.} {Audio-visual tracking of multiple people, using
multiple sensors.
\newline  Member of the Bell South SNT Center, Atlanta, GA.}

\studentempty{Byungmoon Kim}{(CoC)}{Spring 2002 - Present.}

\end{description}


%\student{Alexander Stoychev}{(CoC)}{Fall 1998 - Present}
%\textit{Title: ``Vision Systems for a Robotic Pet.''}
%Jason Brotherton
%\textit{Winter 1999- Present}
%\textit{Title: ``Vision system in C2000.''}
%
%\bigskip
%\centerline{\textbf{College of Computing, Georgia Institute of
%Technology.}}
%\bigskip
%Gabriel Brostow
%\textit{Fall 1997- Present}
%\textit{Admitted to CoC Ph.D. Program Fall 1998}

\subsubsection*{M.S. Thesis Students supervised.}

\begin{description}

\studentnopub{Ken Miller}{(MIT)}{Fall 1994 - Spring
1996.}{Graduated: Spring 1996 \newline Thesis title:
``Eigenmethods for Representing 3D Face Data.''}

\student{Sumit Basu}{(MIT, Co-advised with Dr. Alex
Pentland)}{Fall 1994 - Spring 1996.}{\ref{pub:headtrack-icpr96},
and \ref{pub:faces-ca96}.}{Post Doc at Microsoft Research 2002.}

\student{Arno Sch\"{o}dl}{(CoC)}{Fall 1997 -
Present.}{\ref{pub:headtrack-pui98}, \ref{pub:headtrack-pdpta99}.}
{Completed MS June 1999. Continuing towards a Ph.D.}

\student{Katherine Sukel}{(Psychology, with Prof. Richard
Catrambone)}{Winter 1998 - Present.}{\ref{pub:dance-ijhci00}.}
{Completed MS June 1999. Admitted to Ph.D. Program in CoC.
Currently on Leave.}

\studentnopub{Sunil Mishra}{(CoC)}{Fall 1997 - Fall
1999.}{Research on analyzing audio and video streams to extract
relevant communicative features.\newline Working at SRI in
California.}

\student{Rohit Varma}{(CoC)}{Summer 2000 -
Present.}{\ref{pub:video-matting01}.}{Research on Alpha matting
and analyzing videos of crowds. \newline Currently a PhD student
at Stanford.}

\studentnopub{J Austin Hijar}{(CoC)}{Summer 2001 -
Fall 2001.}{Research on Sound Spaces with Vision Support.}

\studentnopub{Ramprasadh Ramanarayanan}{(CoC)}{Fall 2001 -
Present.}{Research on Fusion of Audio and Video for Interactive Agents.}

\studentnopub{Jeannie Lee}{(CoC)}{Spring 2002 -
Present.}{Research on Fusion of Audio and Video for Distributed Computing.}




%\bigskip

%\centerline{\textbf{Media Laboratory, Massachusetts Institute of
%Technology.}}

%\bigskip



\end{description}

\subsubsection*{MS. Special Problems students.}

\begin{description}

\studentnopub{Junhoy Kim}{(Math)}{Winter 1997 - Summer
1997.}{Title: ``Hidden Markov Models for Vision-based Recognition
of Action.''}

\studentnopub{Alexander Gdalevich}{(ECE)}{Spring 2000 - Fall
2000.}{Research on Embedded Vision Sensors.}

\end{description}

\subsubsection*{Undergraduate Research Students.}

\begin{description}

\studentnopub{Thad Starner}{(MIT)}{Spring 1989 - Spring 1992}
{Title: ``Efficient C implementation of the ThingWorld Modeling
System.'' \newline Finished Ph.D. from MIT and joined Georgia Tech
/ College of Computing as faculty (January 1999).}

%\studentnopub{Roman Khramez}{(CoC)} {Summer 1997 - Spring 1999.}
%{Enrolled as a Ph.D. student in CoC (1999).}

\studentnopub{Roberto Peon}{(CoC)} {Spring 1997 - Spring 2000.}
{Awarded the Georgia Tech Undergraduate Research Internship Award,
1999.
\newline Awarded the Undergraduate Research in Computing Research
Award, 2000.
\newline Currently on staff at Sportvision Inc. }

\studentempty{Charles Hubbard}{(CoC)}{Winter 1998- Fall 1999.}

\studentempty{Paul Bennett}{(CoC)}{Winter 1999-Fall 1999.}

\studentempty{Robbie Walia}{(CoC)}{Fall 2000 - Spring 2001.}

\studentnopub{Tim Keenan}{(CoC)}{Fall 2000 - Spring 2001.}{Awarded
Undergraduate Research in Computing Award 2001. \newline Working
at Rainbow Studios.}

\studentnopub{James Hayes}{(CoC)}{Spring 2001 -
Present.}{\textbf{Awarded Intel Undergraduate Research Fellowship
Award 2001.}
\newline \textbf{Awarded the Georgia Tech College of Computing's Outstanding Undergraduate Research Award 2002.}}

%\studentempty{Stephen Kloder}{(CoC)}{Spring 2001 - Present.}

\studentnopub{Scott Carter}{(CoC)}{Spring 2000 - Present.}{Working
on Digital Video Effects production and support. \newline
Interning with Rhythm and Hues.}

\item \textbf{About 15 other Undergraduate students from 1990 -
Present.}

\end{description}

\subsection{Teaching Honors and Awards}

\begin{enumerate}

\item Georgia Tech. Outstanding use of Innovative Technologies in
Teaching Award, 2000.
\item Georgia Tech. College of Computing's William A. "Gus" Baird Faculty Teaching Award, 2002.

\end{enumerate}

\newpage
\section{RESEARCH AND CREATIVE SCHOLARSHIP}
\label{sec:research}

\subsection{Thesis}
\label{subsec:thesisch}

\begin{description}

\item[S.M. Thesis]{~
\newline Title: {\sl ``Contact Detection, Collision Response and
Friction for Physically-based Virtual World Modeling and Vision
Systems.''}
\newline Date Completed: {\sl June 1990,}
\newline Advisors: {\sl Dr. Alex Pentland and Dr. John Williams,}
\newline University: {\sl Massachusetts Institute of
Technology}. }

\item[Ph.D. Thesis]{~
\newline Title: {\sl ``Analysis, Interpretation, and Synthesis of
Facial Expressions.''}
\newline Date Completed: {\sl February 1995},
\newline Advisor: {\sl Dr. Alex Pentland},
\newline University: {\sl Massachusetts Institute of
Technology}. }

\end{description}

\subsection{Published Journal Papers (refereed)}
\label{subsec:journals}

\begin{pub}

\item Pentland A., I. Essa, M. Friedmann, B. Horowitz, and T. Starner, ``The
ThingWorld Modeling System: Virtual Sculpting by Modal Forces'',
In \textit{ACM Computer Graphics}, 24(2), pp. 143-144, ACM Press,
March 1990.

%Symposium on Interactive 3D Graphics, Snowbird, Utah, pp. 143-144

\item Essa, I., S. Sclaroff, and A. Pentland. ``A Unified Approach
for Physical and Geometric Modeling for Graphics and Animation.''
In \textit{Computer Graphics Forum, The International Journal of
the Eurographics Association}, Vol. 11 (3), C129-C138, C470--C471,
Cambridge, England, September 1992.

\item Darrell, T., I. Essa, and A. Pentland. ``Task-specific Gesture Modeling
using Interpolated Views.'' In\textit{, IEEE Transactions on
Pattern Analysis and Machine Intelligence}, Volume 18 (12), pp.
1236--1242, IEEE Computer Society Press, December 1996.
\label{pub:pami96}

\item Essa. I., and A. Pentland. ``Coding, Analysis, Interpretation and
Recognition of Facial Expressions.'' In\textit{, IEEE Transactions
on Pattern Analysis and Machine Intelligence}, Volume 19 (7), pp.
757--763, IEEE Computer Society Press, July, 1997.
\label{pub:pami97}

\item Essa, I. ``Computers Seeing People'', in \textit{AI Magazine,}
Volume 20 (1), pp. 69--82, AAAI Press, Summer 1999.
\label{pub:aimag00}

\item Essa, I. ``Ubiquitous Sensing for Smart and Aware Environments'',
 In \textit{IEEE Personal Communications, Special Issue on
Networking the Physical World}, Volume 7(5), pp. 47--49, IEEE
Press. October 2000.

\item Sukel, K. E., R. Catrambone, I. Essa, and G. Brostow, ``Presenting movement
in a computer-based dance tutor.'' In \textit{International
Journal of Human-Computer Interaction (IJHCI).} To Appear, Final
Manuscript Submitted August 2001. \label{pub:dance-ijhci00}

\info{ACM SIGGRAPH Proceedings papers
([\ref{pub:vidtex-siggraph00}], [\ref{pub:blur-siggraph01}]) are
considered final archival publications, with resubmission to a
Journal not permitted. Several journal papers are under review.
See~\ref{subsubsec:submitted-journal}}

\end{pub}


\subsection{Published Books and Parts of Books (Refereed)}
\label{subsec:books}

\subsubsection{Chapters in Books}
\label{subsubsec:chapters}

\begin{pub}
\item Essa, I., S. Sclaroff, and A. Pentland. ``Physically-based Modeling for
Graphics and Vision.'' Chapter in \textit{Directions in Geometric
Computing}, pp. 161--196, R. Martin (Editor),
Information-Geometers, UK., 1993.

\item Pentland, A., S. Sclaroff, B. Horowitz, and I. Essa. ``Modal Descriptions
for Modeling, Recognition and Tracking.'' \textit{In
Three-Dimensional Object Recognition Systems}, Vol. I, A K. Jain
and P. J. Flynn (Editors), pp. 423--445, Elsevier Science
Publishers, 1993.

\item Pentland, A., I. Essa, T. Darrell, A. Azarbayejani and S. Sclaroff.
``Visually Guided Animation.'' In \textit{Interactive Computer
Animation}, N. Thalmann and D. Thalmann (Editors), pp. 143--164,
Prentice-Hall, 1996

\item Essa, I. and Alex Pentland, ``Facial Expression Recognition Using Image Motion.'' In \textit{Motion Based
Recognition}, M. Shah and R. Jain (Editors), Chapter 12 , Kluwer
Academic Publishers, Computational Imaging and Vision Series,
1997. \label{pub:mbr97}

\end{pub}

\subsection{Edited Proceedings and Collections}
\label{subsec:edited}

\begin{pub}

\item Proceedings of Second International Conference on Automatic Face and Gesture
Recognition (Program Chair), Killington, VT., USA. IEEE Computer Society
Press, October, 1996.

\item Proceedings of IEEE Workshop on Nonrigid and Articulated Motion, (Program
Co-chair), San Juan, Puerto Rico, IEEE Computer Society Press, June 1997.

%\item Image and Vision Computing Journal, Special Issue on Face and Gesture
%Recognition, (Guest Editor), Elsevier Press, expected printing
%2001.

\end{pub}


\subsection{Conference Presentations}
\label{subsec:conference}

\subsubsection{Invited Keynote Addresses}
\label{subsubsec:invited}

\begin{pub}


\item ``Non-invasive Motion Analysis'', at \textit{International
Symposium on Human Movement Analysis}, Stockholm, Sweden, July 1994.

\end{pub}

\subsubsection{Conference Presentations with Proceedings (refereed and archival)}
\label{subsubsec:conference-archived}

\info{Refereed publications in respected conferences, with
extensive reviewing, and appearing in archival proceedings.}

\HIDE { \info{For the purposes of this vita, refereed means full
paper (not just abstract) reviewed by two or more peers. Review by
an editor is not considered refereed.} }

\begin{pub}

\item Pentland, A., I. Essa, T. Darrell, and S. Sclaroff. ``Visually Guided
Animation.'', \percent{42\%}In \textit{Proceedings of Computer
Animation 1994 Conference,} pp. 129--138, IEEE Computer Society
Press, Geneva, Switzerland, May 1994. \acceptance{40 paper
accepted of 95 submitted (42\%).}

\item Essa, I. and A. Pentland. \percent{18\%}``A Vision System for Observing and Extracting
Facial Action Parameters'', In \textit{Proceedings of IEEE
Computer Vision Pattern Recognition Conference 1994,} pp. 76--83,
IEEE Computer Society Press, Seattle, WA., June 1994.
\acceptance{Accepted for ORAL presentation. 87 (18\%) papers
accepted for ORAL presentation, 77 accepted for POSTER
presentation of 472 submitted, Overall acceptance of 164 out of
472 (35\%).} \label{pub:cvpr94}

\item Essa, I., T. Darrell, and A. Pentland. ``Tracking Facial Motion'', In
\textit{Proceedings of IEEE Nonrigid and Articulated Motion
Workshop 1994}, pp. 36--42, Austin, TX., IEEE Computer Society
Press, November 1994. \label{pub:nram94}

\item Pentland, A., I. Essa, T. Darrell, A. Azarbayejani and S. Sclaroff.
``Visually Guided Interaction and Animation'' \textit{Proceedings
of the Twenty-eighth Asilomar Conference on Signals, Systems \&
Computers 1994,} pp. 1287--1291, Asilomar, CA., November 1994.

\item Darrell, T., I. Essa, and A. Pentland.\percent{33\%} ``Correlation and Interpolation
Networks for Real-time Expression Analysis/Synthesis'', In G.
Tesauro, D. S. Touretzky, and T. K. Leen (Editors), pp. 909--916,
\textit{Advances in Neural Information Processing Systems
(Proceedings of NIPS) 7,} Denver, CO., MIT Press 1995.
\acceptance{Accepted for ORAL presentation. 152 papers accepted of
462 submitted 33\%.} \label{pub:nips95}

\item Essa, I. and A. Pentland.\percent{9\%} ``Facial Expression Recognition using a Dynamic
Model and Motion Energy'', In,\textit{ Proceedings of the IEEE
International Conference on Computer Vision 1995,} pp. 360--367,
IEEE Computer Society Press, Cambridge, MA., May 1995.
\acceptance{Accepted for ORAL presentation. 56 papers accepted for
ORAL presentation of 599 submitted (9\%). Overall acceptance of
159 of 599 (27\%).} \label{pub:iccv95}



\item Basu, S., I. Essa, and A. Pentland.\percent{25\%}``Motion Regularization for Model-based Head
Tracking.'' In \textit{Proceedings of International Conference on
Pattern Recognition}, pp. C8A.3, Vienna, Austria, August 1996.
\acceptance{253 (25\%) papers accepted for ORAL presentation, and
452 (45\%) accepted as POSTERs, out of 1000 submitted papers.
Accepted as an ORAL (25\%).} \label{pub:headtrack-icpr96}

\item Essa, I., S. Basu., T. Darrell, and A. Pentland, ``Modeling, Tracking and Interactive Animation of
Facial Expressions and Head Movements using Input from Video'', In
\textit{Proceedings of Computer Animation 1996 Conference,} pp.
68--79, IEEE Computer Society Press, Geneva, Switzerland, June
1996. \label{pub:faces-ca96}

\item Sch\"{o}dl, A., A. Haro, and I. Essa, \percent{23\%}``Head Tracking using a Textured
Polygonal Model'', In \textit{Proceedings of Perceptual User
Interfaces Workshop}, (held in Conjunction with ACM UIST 1998),
pp. 43--48,  November 1998. \label{pub:headtrack-pui98}
\acceptance{1 of 9 papers accepted for full ORAL presentation, of
over 40 submissions.}

\item Sch\"{o}dl, A., K. Schwan, and I. Essa, \percent{33\%}``Adaptive Parallelization of
Model-based Head Tracking'', In Proceedings of \textit{1999
International Conference on Parallel and Distributed Processing
Techniques and Applications} \textit{(PDPTA'99}) (7 pages), Monte
Carlo Resort, Las Vegas, Nevada, USA, June 1999.
\acceptance{Accepted as a regular research article (RRA),
acceptance rate 33\%.}\label{pub:headtrack-pdpta99}

\item Moore, D., I. Essa, and M. Hayes, \percent{8\%}``Exploiting Human Actions and Object
Context for Recognition Tasks'', In \textit{Proceedings of IEEE
International Conference on Computer Vision 1999 (ICCV'99)}, pp.
80--86, IEEE Computer Society Press, Corfu, Greece, September
1999. \acceptance{Accepted for ORAL presentation. 46 (8\%)
accepted for ORAL presentation out of 575 submitted.  Overall
acceptance was 177 (31\%).} \label{pub:actions-iccv99}

\item Brostow, G., and I. Essa, \percent{8\%}``Motion-based Video Decompositing.'' In
\textit{Proceedings of IEEE International Conference on Computer
Vision 1999 (ICCV'99)}, pp. 8--13, IEEE Computer Society Press,
Corfu, Greece, September
 1999.
\acceptance{Accepted for ORAL presentation. 46 (8\%) accepted for
ORAL presentation out of 575 submitted.  Overall acceptance was
177 (31\%).} \label{pub:layers-iccv99}

\item Kidd. C., G. Abowd, C. Atkeson, I. Essa, B. MacIntyre, E. Mynatt, T.
Starner, ``The Aware Home: A Living Laboratory for Ubiquitous
Computing Research'', In \textit{Proceedings of Second
International Workshop on Cooperative Buildings 1999}, Editors,
Streitz, J. Siegel, V. Hartkopf, S. Konomi, Pittsburgh. LNCS 1670.
Springer: Heidelberg, 1999.

\item Haro, A., I. Essa, and M. Flickner, \percent{47\%} ``Detecting and Tracking Eyes by Using
their Physiological Properties, Dynamics and Appearance'', In
\textit{Proceedings of IEEE Computer Vision and Pattern
Recognition 2000 Conference}, pp. I163--I166. IEEE Computer
Society Press, Hilton Head, SC, June 2000. \acceptance{Accepted
for Poster Presentation.  220 papers accepted of 466 submitted
(47\%).} \label{pub:pupil-cvpr00}

\item Sch\"{o}dl, A., R. Szeliski, D. Salesin, and I. Essa. \percent{19\%} ``Video Textures'', In
\textit{Proceedings of} \textit{ACM SIGGRAPH 2000 Conference}, pp.
489-498, ACM Press,  New Orleans, LA, August 2000. \acceptance{59
papers accepted of 304 submitted (19\%).} \contribution{Initial
idea by Szeliski. Equal intellectual contribution by all.
Implementation and presentation by
Sch\"{o}dl.}\label{pub:vidtex-siggraph00}

\item Mynatt, E., I. Essa, and W. Rogers, \percent{28\%} ``Increasing the Opportunities for Aging
in Place'', In \textit{Proceedings of} \textit{First ACM Universal
Usability 2000} Conference (Inaugural Conference), pp. 65--71,
Arlington, VA, November 2000. \acceptance{Accepted as an ORAL
presentation, 20 (28\%) papers accepted out of 72 submitted. }
\contribution{Equal intellectual contribution and writing by all.
Presentation by Mynatt.}

\item Sch\"{o}dl, A., and I. Essa. \percent{29\%} ``Learning Methods for
Video-based Animation'', In \textit{Advances Neural Information
Processing Systems 13 (Proceedings of NIPS 2000 Conference)},
T.~K.~Leen, T.~.G.~Dietrterich, and V.~Tresp (Editors), MIT Press,
pp. 1002--1008, May 2001. \acceptance{152 papers accepted of 515
submitted (29\%).}\label{pub:vidtex-nips00} \contribution{Equal
intellectual contribution and writing by Essa and Sch\"{o}dl.
Implementation and Presentation by Sch\"{o}dl.}

\item Haro, A., B. Guenter, I. Essa. \percent{39\%}``Real-time, Photorealistic,
Physically Based Rendering of Human Skin Microstructure'' In
\textit{Proceedings of Eurographics Rendering Workshop}, pp.
53--62, London, England June 2001.\label{pub:skin-eg-render01}
\acceptance{29 papers accepted out of 74 submitted (39\%).}
\contribution{Written by Haro and Essa. Equal intellectual
contribution. Implementation and presentation by Haro.}

\item Steedly, D., and I. Essa. \percent{34\%}``Propagation of Innovative Information in Non-Linear
Least-Squares Structure from Motion'' In \textit{Proceedings of
IEEE International Conference on Computer Vision 2001 Conference},
Volume II, pp. 223--229, IEEE Computer Society Press, Vancouver,
BC, Canada, July 2001. \acceptance{205 papers accepted of 596
submitted (34\%).} \label{pub:sfm-iccv01}\contribution{Equal
intellectual contribution and writing by Essa and Steedly.
Implementation and presentation by Steedly.}

\item Brostow, G., and I. Essa. \percent{21\%} ``Image-based Motion Blur for
Stop Motion Animation'' In \textit{Proceedings of ACM SIGGRAPH
2001 Conference}, Los Angeles, CA., USA. August 2001.
\acceptance{65 papers accepted of 300 submitted (21\%)
.}\label{pub:blur-siggraph01} \HIDE{\contribution{Equal
intellectual contribution and writing by Essa and Brostow.
Implementation and presentation by Brostow.}}

\item Sch\"{o}dl, A. and I. Essa, \percent{30\%}``Depth Layers from Occlusions''
To appear in \textit{Proceedings of IEEE Conference on Computer
Vision and Pattern Recognition}, IEEE Computer Society Press,
Kauaui, Hawaii, December 2001. \label{pub:schoedl01a}
\acceptance{Accepted as a poster, overall acceptance rate of
30\%.}

\item Stillman, S. and I. Essa, \percent{29\%}``Towards Reliable Multimodal Sensing in Aware Environments''
To Appear in \textit{Perceptual User Interfaces (PUI 2001)
Workshop} (held in Conjunction with ACM UIST 2001 Conference, ACM
Publishers, November 2001. \label{pub:stillman01a} \acceptance{17
papers accepted for ORAL presentation of 58 submissions. Accepted
as a SHORT ORAL.}

\item Moore, D. and I. Essa, \percent{44\%}``Recognizing Multi-tasking Activities using Context-Free Stochastic Grammar''
In \textit{Proceedings of Workshop on Models versus
Exemplars in Computer Vision}, (held in Conjunction with IEEE
Conference on Computer Vision and Pattern Recognition Conference),
December 2001. \acceptance{8 papers accepted for ORAL presentation
of 18 submissions (44\%). }\label{pub:moore01a}

\item Reveret, L. and I. Essa, ``Visual Coding and Tracking of Speech Related Facial Motions''
In \textit{Proceedings of Cues in Communications
Workshop}, (held in Conjunction with IEEE Conference on Computer
Vision and Pattern Recognition), December 2001.
\label{pub:reveret01a}

\item Sch\"{o}dl, A. and I. Essa, \percent{41\%}``Controlled Animation of Video Sprites'' To Appear in \textit{First ACM Symposium on Computer Animation 2002,} held in conjunction with ACM SIGGRAPH 2002 Conference, San Antonio, TX, USA, July 2002. \acceptance{22 papers accepted for ORAL presentation of 53 submissions (41\%). }\label{pub:schoedl02a}

\item Moore, D. and I. Essa, \percent{26\%}``Recognizing Multitasked Activities from Video using Recognizing Multitasked Activities from Video using Stochastic Context-Free Grammar'' To Appear in \textit{Proceedings of American Association of Artificial Intelligence (AAAI) Conference 2002,} Alberta, Canada, July 2002. \acceptance{121 papers accepted for POSTER presentation
of 469 submissions (26\%). }\label{pub:moore02a}

\item Abowd, G. A. Bobick, I. Essa, E. Mynatt, and W. Rogers ``The Aware Home: Developing Technologies for Successful Aging'' To Appear in \textit{Proceedings of AAAI Workshop and Automation as a Care Giver}, held in conjunction with American Association of Artificial Intelligence (AAAI) Conference 2002, Alberta, Canada, July 2002. \label{pub:abowd02a}

\item Haro, A. and I. Essa, \percent{21\%}``Learning Video Processing by Example'' To Appear in \textit{Proceedings of International Conference on Pattern Recognition 2002,} Quebec City, Canada, August 2002. \acceptance{XX papers accepted for ORAL presentation
of XXX submissions (21\%). }\label{pub:haro02a}

\end{pub}

\subsubsection{Conference Presentations with Proceedings (refereed)}
\label{subsubsec:conference-refereed}

\info{Publication (as a full paper, a short paper or a technical
note) in peer refereed conference/workshop proceedings.}

\begin{pub}

\item Sclaroff, S., I. Essa, and A. Pentland. ``Vision-based Modeling: An
Application of a Unified Approach for Physical and Geometric
Modeling for Graphics and Animation.'' in \textit{Proceedings of
the Eurographics Workshop on Animations and Simulations 1992,} (16
pages) Cambridge, England, September 1992.

\item Essa, I. and A. Pentland. \percent{39\%}``Facial Expression Recognition using Visually
Extracted Facial Action Parameters'', In, \textit{Proceedings of
the International Workshop on Automatic Face and Gesture
Recognition,} pp. 35--40, Zurich, Switzerland, June
1995.\acceptance{Accepted for ORAL presentation. 31 papers
accepted for ORAL presentation of 80 submitted (39\%)}.

\item Brand, M. and I. Essa. ``Causal Analysis for Visual Gesture Understanding'',
in  \textit{Proceedings of AAAI Fall Symposium on Computational
Models for Integrating Language and Vision}, pp. 23--28,
Cambridge, MA., November 1995.

\item Gardner, A. and I. Essa. \percent{34\%}``Prosody Analysis for Speaker Affect
Determination.'' In\textit{ Proceedings of Perceptual User
Interfaces Workshop}, (held in conjunction with ACM UIST 1997
Conference), pp 45--46, Banff, Canada, October 1997.
\acceptance{Accepted for ORAL presentation. 16 (34\%) papers
accepted as ORALs and 22 (47\%) as POSTERs, of 47 submitted.}
\label{pub:affect-pui97}

\item Abowd, G., C. Atkeson, and I. Essa. ``Computational Perception in Future
Computing Environments'', In \textit{Proceedings of Perceptual
User Interfaces Workshop} (held in conjunction with ACM UIST 1997
Conference), pp. 24--25, Banff, Canada, October 1997.
\acceptance{Accepted for POSTER presentation. 16 (34\%) papers
accepted as ORALs and 22 (47\%) as POSTERs, of 47 submitted.}

\item Stillman, S., R. Tanawongsuwan, and I. Essa, ``A System for Tracking and
Recognizing Multiple People with Multiple Cameras'', In
\textit{Proceedings of Second International Conference on Audio-
Vision-based Person Authentication}, April 1999.
\label{pub:tracking-avbpa99}\ienote{Need This}

\item Moore, D., I. Essa, and M. Hayes, ``Object Spaces: Context Management for
Human Activity Recognition'', In \textit{Proceedings of Second
International Conference on Audio- Vision-based Person
Authentication}, April 1999.
\label{pub:actions-avbpa99}\ienote{Need This}

\item Haro, A., I. Essa, and M. Flickner, ``A Non-invasive Computer Vision System
For Reliable Eye Tracking'', In Proceedings of \textit{ACM CHI
2000 Conference}, (Late Breaking Short Paper) pp. 167--168. The
Hague, Netherlands, April 2000. \label{pub:pupil-chi00}

\item Abowd, G., C. Atkeson, A. Bobick. I. Essa, B. MacIntyre, E. Mynatt, T.
Starner, ``Living Laboratories: The Future Computing Environments
Group at Georgia Institute of Technology'', In \textit{Proceedings
of ACM CHI 2000 Conference}, (Organizational Overview) pp.
215--216. The Hague, Netherlands, April 2000.

\end{pub}

\subsubsection{Conference Presentations without Proceedings (abstract refereed)}
\label{subsubsec:mylabel2}

\begin{pub}

\item Friedmann, M., I. Essa, B. Horowitz, S. Sclaroff, T. Starner, and A.
Pentland. ``Distributed ThingWorld'' Live Demonstration of the system at
\textit{G-Tech, ACM SIGGRAPH Conference, '92,} Chicago, IL., August 1992.

\item Pentland, A., I. Essa, M. Friedmann, B. Horowitz, S. Sclaroff, and T.
Starner. ``ThingWorld'' Live Demonstration of the system at \textit{ACM
Computer Graphics, Second Interactive 3D Graphics Symposium,} Snowbird, UT.,
March 1990.

\item Essa, I. ``Ubiquitous Sensing for Smart and Aware Environments'', A Position
paper at \textit{DARPA / NIST / NSF Workshop on Smart
Environments}, Atlanta, GA., July 1999.

\item Essa, I. and L. Reveret ``Machine Perception of Human Activity: Recognizing and
Modeling Facial Expressions and Affect.'', Invited Presentation at
NIPS workshop on Affective Computing. Beckenridge, CO., USA,
December, 2000. \label{pub:affect-nips00}

\end{pub}

\subsection{Other}
\label{subsec:other}

\subsubsection{Submitted Journal Papers}
\label{subsubsec:submitted-journal}

\begin{pub}

%\item Basu, S., I. Essa, and A. Pentland, ``Motion Regularization for Model-based
%Head Tracking.'' Submitted to\textit{ IEEE Transaction on Pattern
%Analysis and Machine Intelligence}, July 1999. \note{Need to take
%this OUT}.

%\item Essa, I and G. Brostow ``Vision and Graphics for Digital Video
%Special Effects Production'', In \textit{International Journal of
%Pattern Recognition and Artificial Intelligence}. (Accepted for
%Publication with Minor Revisions), 2001. \label{pub:dvfx01}

\item Varma, R., A. Sch\"{o}dl, I. Essa. ``Fast Video Matting
using Color Distance Sets'', In preparation for \textit{ACM
Transactions of Computer Graphics}, August 2001.
\label{pub:video-matting01}

\end{pub}

\subsubsection{Submitted Conference Papers}
\label{subsubsec:submitted-conf}

%\begin{pub}

%\item Reveret, L. and I. Essa, ``Characterization of the Manifold of Visual Facial Speech
%Parameters'' Submitted to \textit{Neural Information Processing
%System (NIPS 2001) Conference}, June 2001. \label{pub:reveret01b}

%\end{pub}

\subsubsection{Technical Reports}
\label{subsubsec:technical-reports}

\begin{pub}

\item Tanawongsuwan, R. A. Stoytchev, and I. Essa ``Robust Tracking of People by a
Mobile Robotic Agent'' \textit{Georgia Institute of Technology,
Technical Report \# GIT-GVU-99-19, 1999.} \label{pub:pepe1999a}

\item Xu, J., D. Lipton, I. Essa, M-H. Sung ``Mandatory
Human Participation: A New Scheme for Building Secure Systems.''
\textit{Georgia Institute of Technology, Technical Report \#
GIT-CC-01-09, 2001.} \label{pub:xu01a}

%\item Sch\"{o}dl, A., R. Varma, and I. Essa. ``Video-based Character Animation'' Submitted to \textit{ACM SIGGRAPH
%2001 Conference},  January 2001. \label{pub:vidtex-submitted01}
%
%\item Reveret L.,T. Keenan, and I. Essa. ``Animated Speakers'' Submitted to \textit{ACM SIGGRAPH
%2001 Conference},  January 2001.
%\label{pub:animatedspeakers-submitted01}
%
%\item Haro, A., B. Geunter, S-B. Kang, and I. Essa. ``Geometeric Enhancement of
%Facial Motion Capture'' Submitted to \textit{ACM SIGGRAPH 2001
%Conference},  January 2001. \label{pub:faces-submitted01}

\end{pub}



\subsubsection{Software}
\label{subsubsec:softwaredl}

\begin{pub}

\item ``The ThingWorld System'', a solid modeling, sculpting software developed at
the MIT Media Laboratory (1988-1992). Used in over 25 Research Institutions.
Developed by a team of six people.

\item ``DFace: A Dynamic Facial Modeling and Analysis Package'', a facial analysis
package developed at MIT Media Laboratory (1993-1995). At present ported and
being further developed for Teleconferencing applications by BT (British
Telecom) UK.

\end{pub}

\subsubsection{Video}
\label{subsubsec:videoce}

\begin{pub}

\item Essa, I., T. Darrell, and A. Pentland. ``Modeling and Tracking
Facial Expression.'', in \textit{Standards of Facial Expressions,
SIGGRAPH Video Review} 1995.

\end{pub}

\subsection{Research Proposals and Grants (Principal Investigator)}
\label{subsec:research}


\def\funding#1#2#3#4#5{\item{{{\bf #1}}
\newline Sponsor: {#2}
\newline Investigator(s): {#3}
\newline Amount: {\sl #4}
\newline {\sl #5}}}


\subsubsection*{a.\quad~Approved and Funded}
%\subsubsection{Approved and Funded}
\label{subsubsec:approved}

\begin{enumerate}

\funding{Interaction, Collaboration, and Application Building
within a 3D Virtual Environment.}{DURIP Program, Office of Naval
Research.}{L. Hodges (PI), B. Ribarsky, I. Essa, J. Rossignac and
three other faculty members from the Graphics, Visualization and
Usability Center.}{ \$ 125,928 (Equipment)}{Submitted: September
1996. Funded: January 1997.}

\funding{Building an Intelligent, Adaptive User-Friendly Agent}
{Yamaha Corporation, Japan. Research and Development Division} {A.
Ram and I. Essa.}{\$ 310,000, for 2 years.}{Submitted: December
1996. Funded: January 1997 - March 1999.}

\funding{Digital Imaging Systems for Science and Engineering.}
{Hewlett Packard Corporation.} {R. Schafer, N. Ezquerra, I. Essa,
and T. Barnwell, and six other faculty members from the Graphics,
Visualization and Usability Center (College of Computing) and the
Center for Signal and Image Processing (School of ECE).}{\$
479,000, (Equipment).} {Submitted: May 1997. Funded: June 1997.}

\funding{The Summer Internship Program for Socio-economically
Disadvantaged Under\-gr\-ad\-uates.} {Office of Naval Research.}
{I. Essa.}{\$ 50,000, Annually for 2 years (1998 and
1999).}{Funded: June 1998 - May 2000.}

\funding{Automated Understanding of Captured
Experiences.}{National Science Foundation, Program on Experimental
Systems 1997.}{ I. Essa (PI), C. Atkeson, G. Abowd, and K.
Ramachandran.}{\$ 861,216 for 3 years, (\$ 60,000 matching from
GT).} {Submitted: December 1997.Funded: September 1998.}

\funding{Data-driven Modeling for Real-Time Interaction and
Animation.}{National Science Foundation. CISE Research
Infrastructure Program 1998.}{J. Hodgins (PI), I. Essa, and C.
Atkeson.}{\$ 120,000 (\$ 60,000 Matching from GT) (Equipment).}
{Submitted: July 1998. Funded: February 1999.}

\funding{Augmenting the Capture and Understanding of Everyday
Experiences}{National Science Foundation, CISE Research
Infrastructure Program 1998.}{G. Abowd (PI), C. Atkeson, I. Essa,
T. Starner, \ldots} {\$ 120,000 (\$ 60,000 Matching from GT),
Equipment.} {Submitted: July 1998. Funded: February 1999.}

\funding{Vision-based Transmedia}{Mitsubishi Electric Research
Labs.}{I. Essa.} {\$ 20,000 for 1999-2000.} {Funded through the
GVU Affiliates Program (1/1/199).}

\funding{Ubiquitous Sensing in an Aware Home.}{Microsoft
Research.} {I. Essa.}{\$ 50,000 for 1 year.}{Funded: May 1999.}

\funding{Vision Technology Research}{Microsoft Research.}{I.
Essa.}{\$ 25,000 for 1 year.}{Funded: January 2000.}

\funding{A Spatio-temporal Representation for Analysis and
Modeling of Facial Expressions from Video.}{National Science
Foundation, CAREER Program.}{I. Essa.}{\$ 300,000 for 4 years.}
{Funded: September 2000.}

\funding{Technology First: Teaching Theory in the Context of
Practice}{Hewlett Packard Corporation}{T. Barnwell (ECE) and
several ECE and CoC Faculty.}{$\sim $\$ 800,000 Equipment Grant
(\$ 100,000 for Imaging and Video Effects Classroom)}{Funded:
February 2000.}

\funding{Video Textures.}{Microsoft Research.}{I. Essa.}{\$ 25,000
for 1 year.}{Funded: March 2000.}

\funding{Mechanisms for Securing Emerging Applications.}{National
Science Foundation,  ITR Program 2000.}{M. Ahamad, I. Essa, and H.
Venkateswaran.}{\$ 484,000 for 3 years.} {Funded: November 2000.}

\funding{Human Identification by Movement.}{DARPA HumanID Program
(ISO)}{A. Bobick, I. Essa, and J. Hodgins.}{\$500,000 for 2
years.}{Funded: August 2000.}

\funding{Aware Home Research Initiative.}{A Consortium of
Companies that includes, Intel, MERL, Motorola, Accenture,
Hewl\-ett Packard.}{G, Abowd, A. Bobick, I. Essa. B. Macintyre, B.
Mynatt, T. Starner.}{$\sim$\$445,000 per year.}{Consortium
Established: Summer 2000.}

\funding{The Aware Home: Sustaining the Quality of Life}{ National
Science Foundation, ITR Program 2001 }{G, Abowd, A. Bobick, I.
Essa, B. Mynatt, and W. Rogers (Psych)}{ \$1,600,000 for 5
years}{Funded, July 2001 - June 2006.}

\funding{A Distributed Programming Infrastructure for Integrated
Smart Sensors.}{National Science Foundation ITR 2001 Program}{U.
Ramachandran, S. Deweerth, I. Essa, K. MacKenzie, T.
Starner}{$\sim$ \$1,450,000 for 5 years.}{Funded, July 2001 - June 2006.}

\funding{Mandatory Human Participation: A New Paradigm for
Building Secure Applications}{National Science Foundation, ITR
Program 2001 }{J. Xu, I. Essa, and D. Lipton}{ \$288,000}{Funded, August 2001 - July 2004 .}

\funding{Analysis of Complex Audio-Visual Events Using Spatially Distributed Sensors.}
{National Science Foundation, ITR
Program 2002 }{J. Rehg and I. Essa}{ \$1,000,000}{Approved for Funding,
pending Presidential Approval: May 2002.}

\end{enumerate}

\subsubsection*{b.\quad~Pending}
\label{subsubsec:pendingdentification}
%\begin{enumerate}
%\end{enumerate}

\subsubsection*{c.\quad~Not Funded}
\label{subsubsec:mylabel1}

\begin{enumerate}

\funding{An Interactive System for Pediatric Cardiac Surgery}
{Whitaker Foundation Biomedical Engineering Research Grant 1997}
{I. Essa.}{\$ 208,330 for 3 years}{Submitted: October 1997.}

\funding{Development of a Large Scale Ubiquitous Computing
Interface.}{NSF Major Research Instrumentation Program 1998.}{C.
Atkeson, G. Abowd, I. Essa, and K. Ramachandran.}{\$ 1,984,964 for
3 years.}{Submitted: January 1998.}

\HIDE{ \funding{Supporting Independent Living for Older Adults by
Combining Cognitive Aging, HCI, and Computational Perception}{NSF,
Program in HCI and Universal Access.}{E. Mynatt, I. Essa, and W.
Rogers (Psych).}{\$ 603,187 for 3 years}{Submitted, September,
1999.} }

\funding{Recognition of Group and Crowd Activity from Visual
Information.}{DARPA MURI 2001 Program.}{A. Bobick, I. Essa, E.
Grimsson (MIT), T. Poggio (MIT), T. Darrell (MIT), L. Guibas
(Stanford), D. Koller (Stanford).}{$\sim$\$3,000,000 for 3
years}{Submitted: November 2000. Under Review}

\HIDE{
\item Automated Understanding and Management of Live Experiences
\newline NSF ITR Program 2000 \newline Investigators: G. Abowd, I.
Essa, A. Waibel (CMU), \& J. Wang (CMU) \newline Amount Requested:
\$ 2,679,184 for 3 years \newline Status: Submitted, February 2000
(Pre-Proposal). }

\end{enumerate}

\subsection{Research Proposals and Grants (Contributor)}
\label{subsec:mylabel3}

\subsubsection*{a.\quad~Approved and Funded}
\label{subsubsec:fundedch}

\begin{enumerate}

\funding{Advanced Media-Oriented Systems Research: Ubiquitous
Capture, Interpretation, and Access.}{National Science Foundation
CISE Instrumentation Program}{U. Ramachandran and several CoC
faculty.}{\$ 1,636,194 for 5 years.}{Submitted: November 1998.
Funded: July 1999.}

\end{enumerate}

\subsubsection*{b.\quad~Pending}
\label{subsubsec:pendingd}

\begin{enumerate}
\funding{Rehabilitation Engineering Research Center on
Technologies for Successful Aging}{National Institute for
Disability Rehabilitation Research}{W. Rogers, J. Sanford, E.
Mynatt (Co-Directors), Essa listed as one of the Program
Directors}{\$ 4,500,000, Nov 1, 2001 - Oct 31 2006}{Pending.}

\end{enumerate}


\subsubsection*{c.\quad~Not Funded}
\label{subsubsec:deniedd}

\begin{enumerate}
\funding{Enhancing the Quality of Life with Context-Aware
Computing, Personalized Information Processing, and Secure
Broadband Communications: Research at the Georgia Tech Residential
Laboratory.}{National Science Foundation ITR 1999 Program}{N.
Jayant, and faculty from ECE and CoC.}{\$5,000,000 for 5
years.}{Submitted: March 2000. Not Funded.}

\end{enumerate}

\subsection{Research Honors and Awards}
\label{subsec:mylabel4}

\begin{itemize}

\item Imlay Fellowship, 1996-98.

%\item Georgia Institute of Technology Nominee for Packard Fellowship, 1996.

\item Edenfield Faculty Fellowship 1997.

\item National Science Foundation, CAREER Award, 2000.

\item Georgia Tech. College of Computing, Outstanding Junior
Faculty Research Award, 2000.

\end{itemize}


\newpage
\section{SERVICE}
\label{sec:serviceld}

\subsection{Professional Activities}
\label{subsec:professional}

\subsubsection{Memberships and Activities in Professional Societies}
\label{subsubsec:memberships}

\begin{itemize}

\item Associate Member, Institute of Electrical and Electronics
Engineers (IEEE).

\begin{itemize}
\item IEEE Computer Society.

\item IEEE Computer Society, Technical Committee on Pattern Analysis
and Machine Intelligence.
\end{itemize}

\item Associate Member, Association for Computer Machinery.

\begin{itemize}
\item SIGGRAPH.
\item SIGCHI.
\end{itemize}

\item Member, Sigma-Xi, Tau Beta Pi.
\end{itemize}

\subsubsection{Conference Committee Activities}
\label{subsubsec:mylabel4}

\begin{enumerate}

\item {\bf Invited Panelist}, \textit{NSF Workshop on Facial Modeling and
Animations}, Philadelphia, PA, USA, November 1993.

\item {\bf Member, Program Committee}, \textit{International Conference on
Multi-modal Interaction}, Beijing, China, October 1996.

\item {\bf Chair, Program Committee}, \textit{Second International Conference on
Automatic Face and Gesture Recognition 1996}, Woodstock, VT., USA,
October 1996.

\item {\bf Invited Panelist}, \textit{NSF/DARPA Workshop on Perception of
Action}, Brewster, MA., USA, May 1997

\item {\bf Co-chair, Program Committee}, ``Workshop on Nonrigid and
Articulated Motion'', In Conjunction with \textit{IEEE Computer
Vision and Pattern Recognition Conference 1997}, Puerto-Rico, June
1997.

\item {\bf Member, Program Committee}, ``Workshop on Recent Advances in
Computer Vision'', In Conjunction with \textit{IEEE International
Conference on Computer Vision 1998}, Bombay, India, January 1998.

\item {\bf Member, Program Committee}, International Conference on
Automatic Face and Gesture Recognition 1998, Nara, Japan, April
1998.

\item {\bf Member, Program Committee}, ``Workshop on Interpretation of
Visual Motion'', In Conjunction with \textit{IEEE Computer Vision
and Pattern Recognition Conference 1998}, Santa Barbara, CA., June
1998.

\item {\bf Member, Program Committee}, \textit{IEEE Computer Vision and Pattern
Recognition Conference 1998}, Santa Barbara, CA., June 1998.

\item {\bf Member, Program Committee}, \textit{Computer Animation
Conference 1998}, Philadelphia, PA, June 1998.

\item {\bf Senior Reviewer \& Member, Program Committee}, \textit{ACM
SIGGRAPH 1998 Conference 1998}, Orlando, FL., July 1998.

\item {\bf Member, Program Committee}, \textit{Computer Animation
Conference 1999}, Geneva, Switzerland, June 1999.

\item {\bf Member, Program Committee}, \textit{International Conference on
Automatic Face and Gesture Recognition 2000}, Grenoble, France,
April 2000.

\item {\bf Member, Program Committee}, \textit{Computer Animation
Conference 2000}, Philadelphia, PA, June 2000.

\item {\bf Member, Program Committee}, \textit{IEEE Computer Vision and Pattern
Recognition Conference 2000}, Hilton Head, SC., June 2000.

\item {\bf Member, Program Committee}, ``Workshop on Modeling People''
at the \textit{IEEE Computer Vision and Pattern Recognition
Conference 2000}, Hilton Head, SC., June 2000.

\item{\bf Member, Program Committee}, \textit{IEEE Workshop on Human
Modeling} in Austin, TX, December 2000.

\item  {\bf Member, Program Committee}, \textit{IEEE International
Conference of Computer Vision} in Vancouver, BC, Canada, July
2001.

\item {\bf Member, Program Committee}, ``IEEE Workshop on Detection and
Recognition of Events in Video'' at the \textit{IEEE International
Conference of Computer Vision} in Vancouver, BC, Canada, July
2001.

\item  {\bf Member, Program Committee}, \textit{Computer Animation
Conference 2001}, Korea, June 2001.

\item {\bf Member, Program Committee}, \textit{ACM Workshop on Perceptual User
Interfaces Workshop (held in Conjunction with UIST 2001)},
Orlando, FLorida, October 2001.

\item {\bf Member, Program Committee}, \textit{IEEE Workshop on
Cues in Communications, (held in Conjunction with CVPR 2001)},
Kauai, Hawaii, December 2001.

\item {\bf Member, Program Committee}, \textit{American Association of Artificial Intelligence Conference 2002}, Alberta,
Canada, July 2002.

item {\bf Member, Program Committee}, \textit{First ACM Symposium on Computer Animation 2002}, San Antonio,
TX, USA, July 2002.

\item {\bf Member, Program Committee}, \textit{International
Conference on Pattern Recognition 2002}, Quebec City, Quebec,
Canada, June 2002.

\end{enumerate}

\subsection{Awards}
\label{subsec:awards}

\begin{enumerate}

\item 1999 Dean's Award for Service to the College of Computing, GA
Tech.

\end{enumerate}
\subsection{On-Campus Georgia Tech Committees}
\label{subsec:mylabel1}

\begin{enumerate}

\item Member, Georgia Tech's College of Computing Awards Committee
1997-1998.

\item Member, Georgia Tech's GVU Center Ad-hoc Committee for HCI
Recruiting 1997-1998.

\item Director, Georgia Tech's College of Computing Undergraduate
Summer Internship Program. 1998-2000.

\item Faculty Advisor, Georgia Tech's Pakistan Student
Association. 1999-present.

\item Member, Georgia Tech's College of Computing Faculty
Recruiting Committee (for last few months only), 2000-2001.

\item Member (Elected), Georgia Tech's College of Computing Dean's Advisory
Committee, 2001-2002.

\end{enumerate}

\subsection{Member of Ph.D. Examining Committees}
\label{subsec:member}



\paragraph{Ph.D. Examining Committee -- Georgia Tech.}

\def\phdcomm#1#2#3#4#5{\item{{#1}, {#2}, {#4}.
\newline {\sl Thesis Title: #3}
%\newline {\sl #4,}
\newline {\sl Principal Advisor: #5}}}



\begin{enumerate}

\phdcomm{Fu-Huei Liu}{Department of ECE, College of Engineering,
Georgia Tech.}{``Rate Quality-based Video Coding.''} {December
1996}{Dr. Russell M. Merserau.}

\phdcomm{Sadik Bayrakeri}{Department of ECE, College of
Engineering, Georgia Tech.}{``Scalable Video Coding using
Spatio-Temporal Interpolation.''}{March 1997} {Dr. Russell M.
Merserau.}

\phdcomm{G. Drew Kessler}{College of Computing, Georgia
Tech.}{``Simple Virtual Environment.''} {July 1997} {Dr. Larry
Hodges.}

\phdcomm{Alan Docef}{Department of ECE, College of Engineering,
Georgia Tech.} {``Tele-medicine Applications of Sub-band Image
Coding at Very Low Bit Rates''}{September 1997}{Dr. Mark Smith.}

\phdcomm{Ram Rao}{College of Computing, Georgia
Tech.}{``Multi-modal coding of lip movements.''} {July 1998} {Dr.
Russell M. Merserau.}

\phdcomm{Xue Li}{College of Computing, Georgia Tech.}{``Scalable
and Adaptive Video Multi-cast.''} {July 1998} {Dr. Mostafa Ammar.}

\phdcomm{Clinton Knight} {School of Electrical and Computer
Engineering, Georgia Tech.}{``WWW-based Automatic Testing of
Analog Circuits.''}{March 1999}{Dr. Stephen DeWeerth.}

\phdcomm{Jack Tumblin}{College of Computing, Georgia
Tech.}{``Three Detail-Preserving Contrast Reduction Methods for
Displayed Images.''} {May 1999} {Dr. Gregory Turk.}

\phdcomm{Ara Nefian}{School of Electrical and Computer
Engineering, Georgia Tech.}{``A Hidden Markov Model-Based Approach
for Face Detection and Recognition.''}{August, 1999}{Dr. Monson
Hayes.}

\phdcomm{James O'Brien}{College of Computing, Georgia
Tech.}{``Modeling Brittle Fracture.''}{November, 1999} {Dr.
Jessica Hodgins.}

\phdcomm{Peter Lindstrom}{College of Computing, Georgia
Tech.}{``Image-driven Polygon Simplification and
Optimization.''}{February, 2000} {Dr. Gregory Turk.}

\phdcomm{Charles Wilson}{School of Electrical and Computer
Engineering, Georgia Tech.}{``Neuromorphic Implementations of
Selective Attention.''}{June 2001}{Dr. Stephen DeWeerth.}

\phdcomm{David Stahl}{College of Computing, Georgia
Tech.}{``Bag-of-Particles: An Interactive, Physically-based
Deformable Model.''}{August 2001}{Dr. Norberto Ezquerra.}

\phdcomm{Ron Metoyer}{College of Computing, Georgia
Tech.}{``Combining Character Intelligence and 2D Direction to
Control 3D Animations.''}{August 2001}{Dr. Jessica Hodgins.}

\phdcomm{Victor Zordan}{College of Computing, Georgia Tech.}
{``Motion Capture Simulations that Hit and React.''}{February 2002}{Dr. Jessica Hodgins.}

\phdcomm{Amos Johnson}{School of Electrical and Computer
Engineering, Georgia Tech.}{``Human Identification: A Method for
Gait Recognition Using Static, Activity Specific Body
Parameters.''}{May 2002}{Dr. Aaron Bobick.}

\phdcomm{Huong Quynh Dinh}{College of Computing, Georgia Tech.}
{``Implicit Shapes: Reconstruction and Transformation''}{May 2002}{Dr. Greg Turk.}

\end{enumerate}

\paragraph{Ph.D. Area Exam Committee -- Georgia Tech.}

\def\phdexam#1#2#3#4{\item{{#1}, {#2}, {#3}.
\newline {\sl Principal Advisor: #4}.}}

\begin{enumerate}

\phdexam{David Stahl}{College of Computing, Georgia Tech.}{Fall
1996}{Dr. Norberto Ezquerra.}

\phdexam{Victor Zordon}{College of Computing, Georgia Tech.}{Fall
1997}{Dr. Jessica Hodgins.}

\phdexam{Ron Metoyer}{College of Computing, Georgia Tech.}{Fall
1997}{Dr. Jessica Hodgins.}

\phdexam{Sunil Mishra}{College of Computing, Georgia Tech.}{Fall
1997}{Dr. Chris Atkeson and Dr. Ashwin Ram.}

\phdexam{Anind Dey}{College of Computing, Georgia Tech.}{Spring
1998}{Dr. Gregory Abowd.}

\phdexam{Huong Quynh Dinh}{College of Computing, Georgia
Tech.}{Spring 1999}{Dr. Greg Turk.}

\phdexam{Elliot Moore}{School of Electrical Engineering, Georgia
Tech.}{Spring 2000}{Dr. Mark Clements.}

\phdexam{Eugene Zhgang}{College of Computing, Georgia
Tech.}{Fall 2001}{Dr. Greg Turk.}


\phdexam{Jun Xiao}{College of Computing, Georgia
Tech.}{Spring 2002}{Dr. John Stasko.}


\phdexam{Mark Carlson}{College of Computing, Georgia
Tech.}{Spring 2002}{Dr. Greg Turk.}


\end{enumerate}

\subsection{Consulting, Advisory, and Other External Appointments}
\label{subsec:consultant}

\begin{itemize}
\item{Limited Partner, Midway Airport Partnership, Chicago, IL, USA. 1996 - Present.}
\item{Limited Partner, Yusuf Midway Partnership, Chicago, IL, USA. 1996 - Present.}
\item{Member Local Advisory Board, DVT Sensors Inc. Norcross, GA, USA. 1997 - 1998.}
\item{Consultant, Microsoft Research, Redmond, WA, USA, 2000 - 2001.}
\item{Lead Technical Advisor, Accusense Inc. Atlanta, GA, USA, 1999 - Present.}
\item{Advisor/Consultant, Logic Junction Inc. Atlanta, GA, USA, 2002 - Present.}
\item{Advisor/Consultant, DVT Sensors, Norcross, GA, USA, 2002 - Present.}

\end{itemize}

\section{NATIONAL AND INTERNATIONAL PROFESSIONAL RE\-CO\-GNITION}
\label{sec:national}

\subsection{Invited Conference Session Chair}
\label{subsec:invited}

\begin{enumerate}

\item IEEE Workshop of Nonrigid and Articulated Motion 1994,
Austin, TX.
\item SIGGRAPH 1998, Orlando FL. Session on Facial Animation.

\end{enumerate}

\subsection{Editorial and Reviewer Work for Technical Journals and Publishers}
\label{subsec:editorial}

\begin{enumerate}
\item Reviewer, IEEE International Conference on Computer Vision,
1993, 1995, 2000.

\item Reviewer, IEEE Computer Vision and Pattern Recognition
Conference, 1992, 1993, 1994, 1996, 1997, 1999 and 2000

\item Reviewer, ACM SIGGRAPH 1996, 1997, 1999, 2000.

\item Reviewer, ACM UIST 1999, 2000.

\item Reviewer, International Journal for Computer Vision (IJCV),
Image and Visual Computing Journal (IVCJ), IEEE Transactions on
Pattern Analysis and Machine Intelligence (PAMI), IEEE
Transactions on Robotics and Automation (R\& A), Telepresence,
IEEE Transaction on Multimedia, and several other leading journals
for Computer Vision and Computer Graphics.

\item Guest Editor, Image and Visual Computing Journal, Special
Issue on Face and Gesture Recognition 2001.

\end{enumerate}

\subsection{Patents}

\begin{itemize}
\item Brostow. G., I. Essa "Motion Blur Generator for Stop Motion
Animation." Provisional Patent Application Filed through GTRC
(ID\# 2449), May 2001.
\end{itemize}


\section{OTHER CONTRIBUTIONS}
\label{sec:other}

\subsection{Seminar Presentations (Invited Papers and Talks at Meetings and Symposia)}
\label{subsec:seminar}

\begin{enumerate}

\item ``Nonrigid and Rigid Dynamic Modeling for Animation \&
Graphics'', at \textit{Ecole Normale Superieure, Department de
Mathematiques et Informatique,} Paris, France, (Host: Dr.
Marie-Paul Gascuel), September 1992.

\item ``Physically-based Modeling for Vision, Graphics and
Control.'', at \textit{Academy of Media Arts,} Cologne, Germany,
(Host: Dr. Bernd Girod), September 1992.

\item ``Visualization using Vision-based Modeling'', at \textit{GMD
Super computing Facility, Visualization Group,} Bonn, Germany,
(Host: Dr. Wolfgang Kreuger), September 1992.

\item ``Visual Interpretation of Facial Expre\-ssions'', at
\textit{NSF Work\-shop on Facial Anima\-tions} Phi\-ladelphia,
PA., November 1993.

\item ``Looking at People: Extracting Human Movements.'' at
\textit{International Workshop on Computer Vision and Parallel
Processing 1995}, (Invited Speaker) Islamabad, Pakistan, January
1995.

\item ``Looking at People: Facial Expressions'', Department of Electrical
Engineering Seminar, at \textit{The Beckman Institute, University of Illinois, Urbana-Champaign,} Urbana, IL., (Host: Dr. Thomas Huang), February
1995.

\item ``Looking at People: Expressions and Gestures'', at Microsoft Research,
Seattle WA., (Host: Dr. Matthew Turk), May 1995.

\item ``Control and Estimation in Systems and Environments that Include People'',
Intelligent Engineering Systems Laboratory Colloquium, at \textit{Massachusetts Institute of Technology, Department of Civil and Environmental Engineering,} Cambridge, MA.,
(Host: Dr. Robert Logcher), November 1995.

\item ``Control and Estimation in Systems and Environments that Include People'',
Computer Science Colloquium, at \textit{New York University, Department of Computer Science,} Cambridge, MA., (Host: Dr. Jack Schwartz),
February 1996.

\item ``Perceiving and Understanding Human Actions for Interactive Video
Environments'', at \textit{Electrotechnical Laboratory}, Ibaraki, Japan, March 1996.

\item ``Extracting Perceiving and Modeling of Human Activity '', in \textit{ATR Research Laboratories Workshop on Intelligent Agents.}, (Invited
Speaker) Kyoto, Japan, March 1996.

\item ``Computational Perception of Human Activities in Scenes and Environments'',
at \textit{Graphics, Visualization and Usability Center}, \textit{Georgia Tech} Atlanta, GA, November 1996.

\item ``Computational Perception of Human Activities in Scenes and Environments'',
at \textit{Center for Signal and Image Processing}, \textit{Georgia Tech}. Atlanta, GA, February 1996.

\item ``Intelligent and Aware Environments'', at CHI 1997, Ubiquitous Computing
Workshop, Atlanta, GA., March 1997.

\item ``Importance of Dynamic Models in Perception of Action'', at NSF/DARPA
Workshop on Perception of Action, Brewster, MA, May 1997.

\item ``Ubiquitous Video and Audio'', at \textit{Broadband Telecommunications Center, Georgia Tech.} Atlanta, GA, March, 1999.

\item ``Aware Home'', at \textit{Mitsubishi Electric Research Laboratory}, Cambridge, MA, (Host: Dr. Dick Waters), August 1999.

\item ``Intelligent and Aware Environments'', at Phillips Research, Briarcliff
Manor, NY, (Host: Jim Schmidt), January, 2000.

\item ``Perceiving and Modeling People, Environments and Interactions'', at
Microsoft Research, Seattle, WA (Host: P. Anandan), August 2000.

\item ``Video-based Rendering and Animation'' at \textit{Graphics, Visualization and Usability Center}, \textit{Georgia Tech} Atlanta, GA, November 2001.

\item ``Temporal Reasoning from Video to Temporal Synthesis of Video'' at  \textit{CMU's Robotics Institute Colloquium}, Pittsburgh, PA, February 2002.

\item Over 10 more seminar/colloquia presentations.

\end{enumerate}


\section{PERSONAL DATA}
\label{sec:personal}

\begin{tabular}{@{\hspace{\leftmargini}}ll}
Born: & 11 December 1965, Karachi, Pakistan. \\
Home Address: & 2655 Canna Ridge Circle NE \\
&   Atlanta, GA 30345-1409.\\
Telephone: & (770) 414-8816.\\
Wife: & Shanila Essa; married 30 December 1990.\\
Children: & Gibran Essa; born 7 December 1998, \\
    & Faiz Essa; born 15 April 2000. \\
Citizenship: & USA (Naturalized September 1996).\\
Email: & {\tt irfan@cc.gatech.edu} \\
WWW: & {\tt http://www.cc.gatech.edu/$\sim$irfan}
\end{tabular}
\HIDE{ }


%\end{document}
