\newpage \setcounter{page}{1}
\lfoot[Printed: \today]{Nick Feamster}
\cfoot{\fancyplain{Revised: June 12, 2013}{Representative Publications}}
\rfoot{Page \thepage~of 1}

\begin{center}
{\Large\textbf{Representative Publications}}\\[0.1in] {\large\textbf{Nick
Feamster}}\\
\end{center}

\section*{New Since Tenure}
\begin{enumerate}
\item V. Valancius, B. Ravi, N. Feamster, A. Snoeren
``Quantifying the Benefits of Joint Content and Network Routing''
{\em ACM SIGMETRICS}
Pittsburgh, PA. June 2013.

\item S. Sundaresan, W. de Donato, N. Feamster, R. Teixeira,
S. Crawford, A. Pescape. ``Broadband Internet Performance: A View From
the Gateway'' {\em ACM SIGCOMM} Toronto, Ontario, Canada. August 2011. 

\item V. Valancius, C. Lumezanu, N. Feamster, R. Johari, V. Vazirani
``How Many Tiers? Pricing in the Internet Transit Market''
{\em ACM SIGCOMM}
Toronto, Ontario, Canada. August 2011. 
\end{enumerate}


\section*{Other Representative Publications}
\begin{enumerate}
\item M. Tariq, A. Zeitoun, V. Valancius, N. Feamster, and
M. Ammar. ``Answering What-if Deployment and Configuration Questions
with 'WISE': Techniques and Deployment Experience". {\em IEEE/ACM
Transactions on Networking}, January 2013. 
(Previously appeared in {\em ACM SIGCOMM} Seattle, Washington. August 2008.)

\item B. Anwer, M. Tariq, M. Motiwala, N. Feamster. ``SwitchBlade: A Platform
for Rapid Deployment of Network Protocols on Programmable Hardware''
{\em ACM SIGCOMM}, New Delhi, India, August 2010.

\item S. Hao, N. Syed, N. Feamster, A. Gray and S. Krasser. "Detecting
Spammers with SNARE: Spatio-temporal Network-level Automatic Reputation
Engine". {\em USENIX Security Symposium}, August 2009. Montreal, Quebec,
Canada. 
\end{enumerate}
