\documentclass{article}
\usepackage{fullpage}
\usepackage{bibunits}
\usepackage[ManyBibs]{currvita}

%\textwidth 6.7in
%\oddsidemargin -.11in
%\evensidemargin -.11in
\textwidth 6.6in
\oddsidemargin -.05in
\evensidemargin -.05in
\textheight 9.2in
\topmargin -.1in

\def\up#1{\raise.16ex\hbox{#1}}

\begin{document}
\begin{cv}{}
\textbf{\huge David G. Andersen}\\
\bigskip
MIT Lab for Computer Science \\
NE43-512\\
Cambridge, MA 02139 \\
(617) 452-2821\\
\smallskip
\texttt{dga@lcs.mit.edu}\\
\URL{http://www.angio.net/personal/}

\begin{cvlist}{Education}
  \item \textbf{Massachusetts Institute of Technology}
        \dotfill Cambridge, MA


\sectiontitle{\mdseries {\sc Education}}
\employer{{\sc Massachusetts Institute of Technology}} 
\location{Cambridge, MA}
\verb$    $ M.S./Ph.D. candidate in Computer Science.\\
\verb$    $ {\it Advisor:}  Hari Balakrishnan

\employer{{\sc University of Utah}} \location{Salt Lake City, UT}
\verb$    $ Bachelor of Science in Computer Science.  {\it Cum Laude, 1998}\\
\verb$    $ Bachelor of Science in Biology.  {\it Cum Laude, 1998}

\sectiontitle{\mdseries\sc Research Interests}
Helped implement the Congestion Manager architecture in Linux,
designed and implemented a congestion avoidance scheme for
the {\tt vat} audio tool using the CM.  Developing an architecture
for resilient overlay networks.

\sectiontitle{\mdseries\sc Research Experience}

\sectiontitle{\mdseries\sc Teaching Experience}

\sectiontitle{\mdseries\sc Work Experience}

\sectiontitle{\mdseries\sc Publications}

      The Flask Security Architecture: System Support for Diverse
      Security Policies. Ray Spencer, Stephen Smalley, Peter Loscocco,
      Mike Hibler, David Andersen, Jay Lepreau. Proc. of
      the Eighth USENIX Security Symposium, August 1999.


\sectiontitle{\mdseries\sc Service}

\sectiontitle{\mdseries\sc Other Interests}

\sectiontitle{\mdseries\sc Awards}


\employer{MIT Vinton Hayes Fellowship} \location{1999}
\employer{University of Utah Graduating Student Leadership Award} \location{1998}
\employer{Member, Phi Kappa Phi academic honor society} \location{1995}
\employer{Member, Golden Key academic honor society} \location{1995}
\employer{National Merit Scholar} \location{1993}


\end{llist}

\end{document}
