\documentclass[11pt]{article}

  \usepackage{330-f12}

  \pagestyle{empty}

\begin{document}

  \header{\course}{Parsing}{\Term}

  \vspace{-3mm}

  \begin{enumerate}

    \addtolength{\itemsep}{12mm}

    \item Calculate the first sets for S, A, and B in the following grammar:

          \begin{grammar}[1.35]

            \production{S}{A\mathit{dg} \midspc B\mathit{h}}
              \\

            \production{A}{\mathit{a}B\mathit{c} \midspc \largeepsilon}
              \\

            \production{B}{\mathit{f} \midspc \mathit{g}}
              \\

          \end{grammar}

    \item Rewrite the following grammars so they can be parsed by a
          top--down recursive descent parser by eliminating left recursion
          and applying left factoring where necessary.  (Keep in mind that
          left recurion \textbf{only} needs to be elmininated in a grammar
          if you want to write a top--down recursive descent parser for it.)

          \vspace{-2.5mm}

          \begin{enumerate}

            \addtolength{\itemsep}{4mm}

            \item \production{S}{S + \mathit{a} \midspc \mathit{b}}

            \item \production{S}{S + \mathit{a} \midspc S + \mathit{b} \midspc
                                 \mathit{c}}

            \item \production{S}{\mathit{abc} \midspc \mathit{ac}}

            \item \production{S}{\mathit{aa} \midspc \mathit{ab} \midspc
                                 \mathit{a}}

            \item \production{S}{\mathit{a}S\mathit{c} \midspc
                                 \mathit{a}S\mathit{b} \midspc \mathit{b}}

          \end{enumerate}

          \vspace{-2.5mm}

    \item Consider the following grammar:
          \production{S}{S \ and \ S \midspc S \ or \ S \midspc (S) \midspc
                         true \midspc false}

          \vspace{-2mm}

          \begin{enumerate}

            \addtolength{\itemsep}{4mm}

            \item Give the first sets for each production and nonterminal in
                  the grammar.

            \item Show why the grammar cannot be parsed by a top--down
                  recursive descent parser.

          \end{enumerate}

    \item Consider the following grammar:
          \production{S}{\mathit{ab}S \midspc \mathit{ac}S \midspc
                         \mathit{c}}

          \vspace{-2mm}

          \begin{enumerate}

            \addtolength{\itemsep}{4mm}

            \item Give the first sets for each production and nonterminal in
                  the grammar.

            \item Show why the grammar cannot be parsed by a top--down
                  recursive descent parser.

            \item Rewrite the grammar so it can be parsed by a top--down
                  recursive descent parser.

            \item Write a top--down recursive descent parser for the
                  modified grammar.  You may use the pseudocode notation
                  from the examples in class, and may assume the existencee
                  of functions \texttt{match()} and \texttt{error()}, whose
                  effects are as explained in class.

          \end{enumerate}

  \end{enumerate}

\end{document}
