\documentclass[11pt]{article}

  \usepackage{330-f12}

\begin{document}

  \header{\course}
         {
           \begin{tabular}[t]{@{}c@{}}
             Formal languages and regular expressions -- \\
             solutions
           \end{tabular}
         }
         {\Term}

  \begin{enumerate}

    \addtolength{\itemsep}{6mm}

    \item \begin{enumerate}

            \addtolength{\itemsep}{2mm}

            \item $ \{ \: a, \, aa, \, aaa\: \} \ \: ( \equiv \mathrm{A}) $

            \item $ \{ \: aa, \, aaa, \, aaaa, \, aaaaa, \, aaaaaa \: \} $

            \item $ \{ \: a, \, aa, \, aaa, \, aaaa, \, aaaaa, \,
                       aaaaaa \: \} $

            \item $ \{ \: \epsilon, \, a, \, aa, \, aaa, \, aaaa, \, aaaaa, \,
                       aaaaaa, \, \ldots \: \} $

            \item $ \{ \: bb \: \}  \ \: ( \equiv \mathrm{B})$

            \item $ \{ \: bbbb \: \} $

            \item $ \{ \: bbbbbb \: \} $

            \item $ \{ \: bb, \, bbbb, \, bbbbbb \: \} $

            \item $ \{ \: \epsilon, \, bb, \, bbbb, \, bbbbbb, \, bbbbbbbb, \,
                       bbbbbbbbbb, \, \ldots \: \} $

            \item $ \{ \: abbabb, \, abbaabb, \, abbaaabb, \, aabbabb, \,
                       aabbaabb, \, aabbaaabb, \, aaabbabb, \, aaabbaabb, \,
                       aaabbaaabb \: \} $

            \item $ \{ \: aa, \, aaa, \, aaaa, \, abb, \, aaaaa, \, aabb, \,
                       aaaaaa, \, aaabb, \, bba, \, bbaa, \, bbaaa, \, bbbb
                       \: \} $

          \end{enumerate}

    \item Note that different solutions are possible.

          \newcommand{\sss}{\hspace{.35mm}}
          \newcommand{\mymid}{\mbox{\ensuremath{\sss\mid\sss}}}
          \newcommand{\common}{\ensuremath{( \sss a \mymid b \sss)}}

          \vspace{-1.5mm}

          \begin{enumerate}

            \addtolength{\itemsep}{2mm}

            \item \(
                    abab \, \common^{*}
                  \)
                  ~or~
                  \(
                    abab \, (a^{*}b^{*})^*
                  \)

                  \smallskip

                  The first of these is arguably clearer, but both describe
                  the same language.

            \item \(
                    \common^{*} \, abab
                  \)
                  ~or~
                  \(
                    (a^{*}b^{*})^{*} \, abab
                  \)

            \item \(
                    ( \: ab \, \common^{*} \, ba \: \mid \: aba \: )
                  \)

            \item \(
                    b^*ab^*ab^* \, (\: ab^*ab^*ab^*ab^*ab^* \: )^*
                  \)

            \item \(
                    ( \:
                      b^{*}
                      \, \mid \,
                      ( \: b^{*}ab^{*}ab^{*} \: )^{*}
                      \, \mid \, 
                      ( \: \common \: \common \: )^{*} \:
                    )
                  \)
                  ~or~
                  \(
                    ( \:
                      b^{*}
                       \, \mid \,
                      ( b^{*}ab^{*}ab^{*} )^{*}
                       \, \mid \,
                      ( \: ab \mid bb \mid ba \mid aa \: )^{*}  \:
                    )
                  \)

            \item \(
                    \common^{*} \ aaa \ \common^{*}
                  \)

            \item \(
                    b^{*} \, ( \, ( \epsilon \mid a \mid aa \, ) \, b \, )^* \
                    ( \epsilon \mid a \mid aa )
                  \)

          \end{enumerate}

    \item One thing which is interesting which you may have noticed is that
          even making a small change in the regular expression can have a
          big effect on the results it produces.

          \vspace{-1.5mm}

          \begin{enumerate}

            \addtolength{\itemsep}{2mm}

            \item This regular expression is wrong.  The two strings 1 and
                  11 for example are in the language, so the regular
                  expression should recognize or accept them, but it does
                  not.

            \item This regular expression is wrong.  The string 1001 for
                  example is in the language, so the regular expression
                  should recognize or accept it, but it does not.  The same
                  would apply to any other valid string which begins or ends
                  with a 1.

            \item This regular expression is wrong.  The string 111 for
                  example is not in the language, so the regular expression
                  should not recognize or accept it, but it does.  The same
                  would apply to any other invalid string containing three
                  or more consecutive 1s.

            \item This regular expression is wrong.  The string 001110 for
                  example is not in the language, so the regular expression
                  should not recognize or accept it, but it does.  The same
                  would apply to some other invalid strings containing three
                  or four consecutive 1s, since the regular expression will
                  recognize or accept strings in which alternate
                  subsequences of 1s may have three or four consecutive 1s.

            \item This regular expression is wrong.  The string 0011 for
                  example is in the language, so the regular expression
                  should recognize or accept it, but it does not.  The same
                  would apply to any other valid string which contains more
                  than zero 0s and which ends with a 1.

            \item This regular expression is wrong.  The string 010100 for
                  example is in the language, so the regular expression
                  should recognize or accept it, but it does not.  The same
                  would apply to other valid strings, since this regular
                  expression requires that every alternate 0 must be
                  followed by a 0, but not every valid string has that
                  property.

            \item This regular expression is correct.

            \item This regular expression is also correct.

          \end{enumerate}

  \end{enumerate}

\end{document}
