
\vspace{.1in}

%\noindent
\subsection*{SYNERGISTIC ACTIVITIES} 
\begin{itemize}
\item Co-organizer, {\em Workshop on Internet Routing Evolution and
  Design (WIRED)}, October 2006.

\item Program Committee Co-Chair, {\em Workshop on the Economics of
  Networked Systems} (formerly {\em P2PEcon} and {\em SIGCOMM PINS
  Workshop}), June 2006.

\item Frequent presenter at North American Network Operators Group.
  Presentations on many topics, including traffic engineering, routing
  security, spam, network configuration and management, and Internet
  address space allocation.

\item Program Committee member, {\em ACM SIGCOMM Internet Measurement
  Conference}, October 2006.

\item Program Committee member, {\em IEEE Symposium on Security and
  Privacy}, May 2006.
\end{itemize}

\vspace{.1in}

%\noindent
\subsection*{COLLABORATORS} 

\begin{itemize}
\item
Advisor: Prof. Hari Balakrishnan (MIT)

%\item
%Advisees: A. Ramachandran (Georgia Tech)

\item
Recent and current collaborators:
D. Andersen (CMU),
A. Bavier (Princeton),
R. Bush (IIJ),
M. Caesar (UC Berkeley),
D. Caldwell (AT\&T),
L. Gao (UMass Amherst),
M. Huang (Princeton),
R. Johari (Stanford), 
W. Lee (Georgia Tech),
J. Lepreau (Utah),
Z. Mao (U. Michigan),
D. Mazieres (Stanford),
L. Peterson (Princeton),
J. Rexford (Princeton),
S. Shenker (UC Berkeley),
K. van der Merwe (AT\&T),
V. Vazirani (Georgia Tech), 
J. Winick (Lockheed Martin), 
J. Xu (Georgia Tech)
\end{itemize}



\end{document}

\pagebreak

\section{Budget Justification}


The domestic travel allowance is for attending three domestic
conferences, for instance, SIGCOMM, NSDI, ACM SIGCOMM Internet
Measurement Conference, and making trips for research collaborations
(such as ICSI, Berkeley).  The international travel allowance is for
attending these conferences when they are held outside the U.S.; many of
the ACM SIGCOMM conferences are to be held outside of the U.S. in the
coming years.
