\documentclass[11pt]{article}
%\documentclass[extras,11pt]{article}


\setlength{\textheight}{9.2in}
\setlength{\textwidth}{6.8in}
%\setlength{\topmargin}{-0.2in}
\setlength{\evensidemargin}{0pt}
\setlength{\oddsidemargin}{0pt}
\setlength{\headsep}{10pt}
\setlength{\parskip}{1ex}



\begin{document}


\section*{Biographical Sketch: Nick Feamster}



\subsection*{EDUCATION}

%\vspace{1ex}
\begin{tabbing}
\ \ \ \ \ \=Spring 2006-\ \ \ \=Research Advisor: Prof. Manuel Blum\ \ \=   \kill
\>2000\> S.B., Electrical Engineering \> Massachusetts Institute of
Technology\\ \>\> \ \ \ \ \ and Computer Science \> \\

\>2001\> M.Eng., Computer Science \> Massachusetts Institute of Technology\\
\>2005\> Ph.D., Computer Science \> Massachusetts Institute of Technology\\
\> \>{\em Research Advisor:} Prof. Hari Balakrishnan \\
\end{tabbing}

\subsection*{APPOINTMENTS}

%\vspace{1ex}
\begin{tabbing}
\ \ \ \ \ \=Spring 2006-\ \ \ \=Research Advisor: Prof. Manuel Blum\ \ \=   \kill

\> Spring 2006-\> Assistant Professor \> Georgia Institute of Technology \\
\> \> \> College of Computing, Atlanta, GA  \\
\> Fall 2005\> Post-Doctoral Research Staff \> Princeton University,
Princeton, NJ \\
\end{tabbing}

\subsection*{AWARDS}

%\vspace{1ex}
\begin{tabbing}
\ \ \ \ \ \= \kill
\> Sloan Research Fellow \\
\> IBM Faculty Research Award \\
\> NSF CAREER Award \\
\> Advisor for Best Student Paper, {\em ACM SIGCOMM 2006} \\
\> Best Paper, {\em 2nd Usenix Symposium on Networked Systems Design
  and Implementation} \\
%\> Cisco URP Grant Recipient \\
\> Best Student Paper, {\em 11th Usenix Security Symposium} \\
\> Best Student Paper, {\em 10th Usenix Security Symposium} \\
\> MIT William A. Martin Memorial Thesis Award for best EECS Master's
thesis \\
\> NSF Graduate Research Fellow \\
\end{tabbing}



%\noindent
\subsection*{FIVE RELATED PUBLICATIONS} 

\begin{enumerate}

\item
Anirudh Ramachandran, Nick Feamster, and Santosh Vempala
{ Filtering Spam with Behavioral Blacklisting }.
In {\em {Proc.\ ACM Conference on Computer and Communication Security
(CCS)}}, Alexandria, {VA}, November 2007. 

\item 
David G. Andersen, Hari Balakrishnan, Nick Feamster, and Scott
Shenker. Holding the Internet Accountable. In {\em Proc. 6th ACM
  Workshop on Hot Topics in Networks (Hotnets-VI)}, Atlanta, GA,
November 2007.

\item
Anirudh Ramachandran and Nick Feamster
{Understanding the Network-Level Behavior of Spammers}.
In {\em {Proc. {ACM SIGCOMM}}}, pages 291--302, Pisa, Italy, September 2006.

\item
Yiyi Huang, Nick Feamster, Anukool Lakhina, and Jim Xu
{Diagnosing Network Disruptions with Network-Wide Analysis}.
In {\em {Proc. ACM SIGMETRICS}}, pages 61--72, San Diego, CA, June 2007.


\item
Nick Feamster, David Andersen, Hari Balakrishnan, and M.~Frans Kaashoek.
{Measuring the Effects of {I}nternet Path Faults on Reactive Routing}.
In {\em {Proc. ACM SIGMETRICS}}, pages 126--137, San Diego, CA, June 2003.


\end{enumerate}


%\noindent
\subsection*{FIVE SIGNIFICANT PUBLICATIONS} 
\begin{enumerate}



\item
Andy Bavier, Nick Feamster, Mark Huang, Larry Peterson, and Jennifer Rexford.
{In VINI Veritas: Realistic and Controlled Network Experimentation}.
In {\em {Proc. {ACM SIGCOMM}}}, pages 3--14, Pisa, Italy, September 2006.


\item
Nick Feamster, Ramesh Johari, and Hari Balakrishnan.
{The Implications of Autonomy for Stable Policy Routing}.
In {\em {Proc.\ ACM SIGCOMM}}, pages 25--36, Philadelphia, {PA}, August 2005.

\item
Nick Feamster and Hari Balakrishnan.
{Detecting BGP Configuration Faults with Static Analysis}.
In {\em Proc. 2nd Symposium on Networked Systems Design and
  Implementation (NSDI)}, pages 43--56, Boston, MA, May 2005.

\item
Matthew Caesar, Don Caldwell, Nick Feamster, Jennifer Rexford, Aman
Shaikh, and Kobus van~der Merwe.
{Design and Implementation of a Routing Control Platform}.
In {\em Proc. 2nd Symposium on Networked Systems Design and Implementation (NSDI)}, pages 15--28, Boston, MA, May 2005.

\item
Nick Feamster, Jared Winick, and Jennifer Rexford.
{A Model of {BGP} Routing for Network Engineering}.
In {\em {Proc. ACM SIGMETRICS}}, pages 331--342, New York, NY, June 2004.
Also in {\em IEEE/ACM Transactions on Networking}.

%% \item 
%% Nick Feamster, Jay Borkenhagen, and Jennifer Rexford.
%% {Guidelines for Interdomain Traffic Engineering}.
%% {\em {ACM Computer Communications Review}}, 33(5):19--30, October 2003.


%\item
%Nick Feamster, Magdalena Balazinska, Greg Harfst, Hari Balakrishnan, and
%David Karger. 
%Infranet: Circumventing {W}eb censorship and surveillance.
%In {\em {Proc. 11th USENIX Security Symposium}}, {San Francisco, CA}, August 2002.

%% \item
%% Kevin Fu, Emil Sit, Kendra Smith, and Nick Feamster.
%% Dos and don'ts of client authentication on the {W}eb.
%% In {\em {Proc. 10th USENIX Security Symposium}}, {Washington, DC}, August 2001.


\end{enumerate}

%\noindent
\subsection*{SYNERGISTIC ACTIVITIES} 
\begin{itemize}
\item Frequent presenter at North American Network Operators Group.
  Presentations on many topics, including traffic engineering, routing
  security, spam, network configuration and management, and Internet
  address space allocation.

\item Program Committee, {\em ACM SIGCOMM}, August 2008.

\item Program Committee, {\em ACM SIGMETRICS}, June 2008.

\item Program Committee, {\em WWW Conference}, Security and
  Privacy Track, May 2007, May 2008.

\item Program Committee, {\em ACM SIGCOMM Internet Measurement
  Conference}, October 2006.

\item Program Committee, {\em IEEE Symposium on Security and
  Privacy}, May 2006.

\item Co-organizer, {\em Workshop on Internet Routing Evolution and
  Design (WIRED)}, October 2006.

%\item Program Committee Co-Chair, {\em Workshop on the Economics of
%  Networked Systems} (formerly {\em P2PEcon} and {\em SIGCOMM PINS
%  Workshop}), June 2006.

\end{itemize}

%\noindent
\subsection*{COLLABORATORS} 

\begin{itemize}
\itemsep=-1pt
\item
Advisor: Prof. Hari Balakrishnan (MIT)

\item
Ph.D. Advisees: A. Ramachandran, M. Motiwala, Y. Huang, V. Valancius,
M. bin Tariq, N. Syed, S. Hao

\item
Recent and current collaborators:
D. Andersen (CMU),
A. Bavier (Princeton),
R. Bush (IIJ),
M. Caesar (UC Berkeley),
D. Caldwell (AT\&T),
J. Chase (Duke),
C. Diot (Thomson)
L. Gao (UMass Amherst),
A. Gray (Georgia Tech),
C. Guestrin (CMU),
J. Hellerstein (UC Berkeley),
M. Huang (Princeton),
R. Johari (Stanford), 
W. Lee (Georgia Tech),
J. Lepreau (Utah),
Z. Mao (U. Michigan),
L. Peterson (Princeton),
J. Rexford (Princeton),
S. Shenker (UC Berkeley),
R. Teixeira (LIP6)
K. van der Merwe (AT\&T),
V. Vazirani (Georgia Tech), 
S. Vempala (Georgia Tech),
J. Winick (Lockheed Martin), 
J. Xu (Georgia Tech)
\end{itemize}



\end{document}
